---
title: "Comparison of Bayesian, Genearlized Mixed Models and Randomization Approaches"
author: "Ben Weinstein"
date: "Thursday, February 12, 2015"
output: html_document
---

#Aim
  * Ensure that our result could rise arise solely due to the topology of the tree. Since all species are related, there is a finite number of possible curves that could be produced from our observed phylogeny. The goal is to randomize the species tips and create communities equal to the observed richness, but random with respect to phylo-relatedness. If our observed function is outside the $\alpha=.05$ confidence interval, we can conclude that our result is highly unlikely to be influenced by the topology of our tree.
  
  * Compare the estimates from a bayesian model used in the main text with an simliarly formulated model fitted use generalized linear mixed models.


```{r,message=FALSE,warning=FALSE}
library(knitr)
library(pez)
library(ape)
library(dplyr)
library(ggplot2)
library(reshape2)
library(boot)
library(vegan)
library(RColorBrewer)
library(gridExtra)
library(scales)
library(stringr)
library(picante)
library(foreach)
library(doSNOW)
library(lme4)
opts_chunk$set(warning=FALSE,message=FALSE,echo=TRUE,eval=T)
opts_chunk$set(cache=TRUE, cache.path = 'Independence_cache/', fig.path='figure/',fig.width=11,fig.height=6)

#load("Independence.Rdata")

```


```{r,echo=FALSE}
#Define co-occur function

co_occur<-function(dat,trx,ctrx){
 
f<-split(dat,dat$site)

#remove the localities with only 2 species
f<-f[sapply(f,function(x) sum(x$Y)) > 1]

if(length(f)==0){return(NA)}

#Get closest related species in each site
closest<-lapply(f,function(x){
  
  #species present
  pres<-x$sp[x$Y==1]
    dis<-ctrx[rownames(ctrx) %in% trx$tip.label,colnames(ctrx) %in% pres]
    apply(dis,1,function(y){
      as.matrix(min(y[!y==0]))
      })
  }) 

#melt each list, a bit ugly
tomerge<-lapply(closest,function(x){
 dm<-melt(as.matrix(x))[,-2]
 colnames(dm)<-c("sp","x")
 return(dm)})

tomerge<-melt(tomerge,id.var=c("sp"))

tomerge<-dcast(tomerge,...~variable)
colnames(tomerge)<-c("sp","site","x")

#need to be characters not factors
dat$site<-as.character(dat$site)
tomerge$site<-as.character(tomerge$site)

PA_phylo<-merge(dat,tomerge,by=c("site","sp"))
PA_phylo$Y<-as.numeric(PA_phylo$Y)

#remove Inf lengths, no co-occurrence
PA_phylo[!is.finite(PA_phylo$Y),"Y"]<-NA

return(PA_phylo)
}

```

```{r}

simL<-function(species,sites){

#Simulate a balanced phylogeny
trx<-compute.brlen(stree(species,"balanced",tip.label=1:species))

#Cophenetic matrix
ctrx<-cophenetic(trx)

#Create data.frame
sp=rep(1:species,sites)
site=as.vector(sapply(1:sites,function(x) rep(x,species)))
dat<-data.frame(sp,site)

#Presence/absence
dat$Y<-rbinom(sites*species,1,.5)

#Compute co-occurrence
dat<-co_occur(dat,trx,ctrx)
return(dat)}
```


##Observed versus null distribution of covariate estimates

```{r}
##read in data
#read in tree
trx<-read.tree("InputData\\hum294.tre")

new<-str_extract(trx$tip.label,"(\\w+).(\\w+)")
#get duplicates
trx<-drop.tip(trx,trx$tip.label[duplicated(new)])

#name tips.
trx$tip.label<-str_extract(trx$tip.label,"(\\w+).(\\w+)")

ctrx<-cophenetic(trx)

#standardize the distances, just to avoid rounding error.
ctrx<-ctrx/max(ctrx)

siteXspp<-read.csv("ObservedData.csv",row.names=1)
source("SpeciesOverlapSourceFunctions.R")

```

```{r}
#definite function to randomize assemblage and compute glm
randomCo<-function(siteXspp){
  siteXspp$Phylo.RelatednessR<-sample(siteXspp$Phylo.Relatedness)
  mod<-glm(data=siteXspp,formula = P_A ~ poly(Phylo.RelatednessR,2,raw=TRUE), family = "binomial") 
  mode<-summary(mod)$coefficients[,"Estimate"]
  data.frame(Intercept=mode[[1]],x=mode[[2]],x2=mode[[3]])
}
```

```{r,eval=TRUE}
cl<-makeCluster(8,"SOCK")
registerDoSNOW(cl)
out<-foreach(x=1:1000,.packages=c("picante","reshape2")) %dopar% {
  randomCo(siteXspp)
  }

stopCluster(cl)
```

```{r}
#melt data frame
dat<-melt(out)

#remove last column
dat<-dat[,-3] 
```

###Observed Relationship using glm and glmm

What is our observed relationship between probability of presence and phylogenetic distaance to a co-occurring species. I want to both compare the bayesian estimate to a null distrubiton of randomized assemblages, as well as to other modeling approaches.

  * Fit a glm without a random effect
  * Fit a mixed model (glmer) with a species level random effect for intercept, linear and polynomial covariates.

###GLM

```{r model fit,fig.height=4,fig.width=4,cache=TRUE,eval=TRUE}
 

#glm observed
mod<-glm(data=siteXspp,formula = as.factor(P_A) ~ poly(Phylo.Relatedness,2,raw=TRUE), family = "binomial") 
summary(mod)
mode<-summary(mod)$coefficients[,"Estimate"]
observed<-melt(data.frame(Intercept=mode[[1]],x=mode[[2]],x2=mode[[3]]))
#confint.glm<-confint(mod)
predict.glm<-data.frame(x=siteXspp$Phylo.Relatedness,y=predict(mod))
```

### GLMM with random effect on intercept and slopes

```{r}
mod2<-glmer(data=siteXspp,formula = as.factor(P_A) ~ 1 + poly(Phylo.Relatedness,2,raw=TRUE) + (1 + poly(Phylo.Relatedness,2,raw=TRUE)|Species), family = "binomial")
smod2<-summary(mod2)
smod2
coef.glmer<-smod2$coefficients
observed.glmer<-melt(data.frame(Intercept=coef.glmer[1,1],x=coef.glmer[2,1],x2=coef.glmer[3,1]))
#confint.glmer<-confint(mod2)
predict.glmer<-data.frame(x=siteXspp$Phylo.Relatedness,y=predict(mod2))

observed.bayes<-melt(data.frame(Intercept=-1.96,x=6.64,x2=-11.69))

#bind results
coef.dat<-melt(list(GLM=observed,GLMM=observed.glmer,Bayesian=observed.bayes)) 
```

#Visualize null distribution versus observed value

  * View the mean estimates against the null distribution
  
```{r,eval=TRUE}

quants<-group_by(dat,variable) %>% summarize(lower=quantile(value,0.025),upper=quantile(value,0.975))

ggplot(dat,aes(x=value)) + geom_histogram() + facet_wrap(~variable,scales="free") + theme_bw() geom_vline(data=coef.dat,aes(xintercept=value,col=L1),linetype='dashed',size=1.1,show_guide = T) + geom_vline(data=quants,aes(xintercept=lower),col='black',size=.7)+ geom_vline(data=quants,aes(xintercept=upper),col='black',size=.7) + labs(col="Modeling Approach") 
  
```

The histogram is the distribution of glm estimates for randomly draw assembalges. We compare this null distrubtion to the observed mean covariate estimates for glm glmm, and bayesian models. 
From this figure I infer that the the probability of getting our result by the intertwined structure of the model is very low. In addition, our result is even more extreme since we need to get both the x and x^2 terms, they are not independent, so the joint probability of getting our result is even more remote than just $\alpha=0.05$ would suggest.

```{r,cache=TRUE}

#define trajectory function
trajF<-function(intercept,linear,polynomial,x){
  p<-inv.logit(intercept + linear * x  + polynomial * x^2)
  return(p)
}

s<-seq(0,1,0.01)

#Simulated trajectories
ynew<-lapply(out,function(x){
  y<-trajF(x$Intercept,x$x,x$x2,s)
  data.frame(s,y)
  })
 
ynew<-rbind_all(ynew)

conf<-group_by(ynew,s) %>% summarise(mean=mean(y),upper=quantile(y,0.975),lower=quantile(y,0.025))


#Observed glm
yobs.glm<-data.frame(x=siteXspp$Phylo.Relatedness,y=trajF(observed[1,2],observed[2,2],observed[3,2],siteXspp$Phylo.Relatedness))

#observed glmer (1|Species)
yobs.glmer<-data.frame(x=siteXspp$Phylo.Relatedness,y=trajF(observed.glmer[1,2],observed.glmer[2,2],observed.glmer[3,2],siteXspp$Phylo.Relatedness))

#observed bayesian with species level effect
confint.bayes<-data.frame("2.5%"=c(-2.44,4.72,-9.37),"97.5%"=c(-1.48,8.62,-14.71))
#colnames(confint.bayes)<-colnames(confint.glm)

yobs.bayes<-data.frame(x=siteXspp$Phylo.Relatedness,y=trajF(observed.bayes[1,2],observed.bayes[2,2],observed.bayes[3,2],siteXspp$Phylo.Relatedness))

obs<-melt(list(GLM=yobs.glm,Mixed_Effects=yobs.glmer,Bayes=yobs.bayes),id.var=c("x","y"))

ggplot(data=conf,aes(x=s,y=mean)) + geom_ribbon(aes(ymin=lower,ymax=upper),fill='gray80') + theme_bw()  + labs(x="Phylogenetic distance to the closest related species",y="Probability of Presence") + geom_line(data=obs,aes(x=x,y=y,col=L1),show_guide=TRUE,size=1.5,linetype="dashed") + ggtitle("Observed versus null fit") + geom_line(aes(y=mean)) + labs(col="Modeling Approach") 

```

The grey area is the null distrubition of glm estimates. The red line is the Bayesian estimate, blue is the glm mixed effects with a random effect for all coefficients, and the green is glm without accounting for species differences. Regardless of model choice, the observed relationship is quite different than the flat relationship we expect based on the topology of the tree.

## Compare bayesian credible intervals with generalized linear models for each on the simulated, predicted, and observed assemblages.

```{r,warning=FALSE,message=FALSE}
#Load required libraries
library(reshape2)
require(ggplot2)
library(picante)
library(dismo)
library(ape)
library(doSNOW)
library(gdistance)
library(foreach)
library(boot)
library(maptools)
library(rasterVis)
library(knitr)
library(vegan)
library(gridExtra)
library(R2jags)
library(stringr)
library(dplyr)
library(scales)
library(lme4)
opts_chunk$set(warning=FALSE,message=FALSE)
opts_chunk$set(cache=TRUE, cache.path = 'SpeciesOverlap_cache/', fig.path='figure/',fig.width=10,echo=TRUE)

#Set git path
gitpath<-"C:/Users/Ben/Documents/Pred_Obs/"
droppath<-"C:/Users/Ben/Dropbox/"

setwd(gitpath)

#Load image if desired
#load(paste(droppath,"Thesis\\Pred_Realized\\Assemblages\\Threshold0.05\\Results\\","PredictedRealized_Revis.RData",sep=""))

```

```{r}
PA_m2<-read.csv("InputData/SpeciesData.csv")
```

##Combine simulation, predicted and observed assemblages.

Read in simulated data - see Simulations.Rmd

```{r}
simall<-read.csv("C:/Users/Ben/Documents/Pred_Obs/Bayesian/simdats.csv",row.names=1)

simall$P_R<-"Simulation"
```

Bind data to one dataframe.

```{r}

#make formats the same
PA_m2$Locality<-as.numeric(as.character(PA_m2$Locality))
PA_m2$P_A<-as.integer(PA_m2$P_A)

#bind with simulated data
PA_m2<-rbind_all(list(PA_m2,simall))

#fill in iteration for non simulations
PA_m2$Iteration[is.na(PA_m2$Iteration)]<-1
#split data into types of assemblages

#drop Env + Dispersal for observed, its just a data subset
sdat<-split(PA_m2,list(PA_m2$P_R,PA_m2$Hyp),drop=TRUE)

names(sdat)<-c("Env","Observed","Env + Dispersal","No Signal & Strong Repulsion","Strong Signal & Strong Repulsion","No Signal & No Repulsion","Strong Signal & No Repulsion")

# remove inf values
sdat<-lapply(sdat,function(y){
    y<-y[is.finite(y$Phylo.Relatedness),]
})

```

On average, how many data points in each simulation and observed data?
```{r}
sapply(sdat,nrow)
```

#Compare to generalized linear models

While these are constrained to have the same intercept and slope for each species, they provide a reasonable check on whether the bayesian estimates are giving logical values.

```{r,fig.height=10}
plotlist<-lapply(sdat,function(x){
p<-ggplot(x,aes(x=Phylo.Relatedness,y=P_A)) + geom_line(aes(group=Iteration),stat="smooth",method="glm",family="binomial",formula=y~poly(x,2),alpha=min(c(1,(1/max(x$Iteration)*10)))) + theme_bw() + labs("x=Distance to closest related species in an assemblage",y="Probability of presence") + ggtitle("") + scale_alpha(guide="none")
return(p)
})

#title plots
for (x in 1:length(plotlist)){
  plotlist[[x]]<-plotlist[[x]]+ggtitle(names(plotlist)[x])
}

do.call(grid.arrange,plotlist)
```

##Define model function

Define a function that takes in species, presence/absence and distance to the closest related species for each type of assemblage, and returns the posterior estimates of X and X^2.

```{r singlemodel,echo=TRUE,eval=TRUE}

GLMM<-function(dat){
mod<-glmer(data=dat,formula = as.factor(P_A) ~ 1 + poly(Phylo.Relatedness,2,raw=TRUE) + (1|Species), family = "binomial")
return(mod)
}
```

##Fit GLMM function for each assemblage type

```{r run jags,results='hide',eval=T}
parall<-lapply(sdat,function(x){
  out<-GLMM(x)
})
```

The next goal is to extract the fixed effects from the GLMM and visualize them.

```{r}
formatfixed<-function(x){

#fixed estimates
f<-fixef(x)
out<-melt(data.frame(Intercept=f[1],Linear=f[[2]],Polynomial=f[[3]]))
out$Var2<-"Mean"

#Find confidence intervals of fixed effects (assuming distribution...)
conf<-confint.merMod(x,method = "Wald")
rownames(conf)<-c("Intercept","Linear","Polynomial")
colnames(conf)<-c("Lower","Upper")
conf<-melt(conf)

#match column names
colnames(out)<-c("Var1","value","Var2")

#bind mean and confidence intervals
out<-rbind(out,conf)
return(out)
}
```

```{r}
#get fixed effects
fixf<-lapply(parall,formatfixed)
```

Plot fixed effects against underlying data

```{r}
#define trajectory function
trajF<-function(intercept,linear,polynomial,x){
   p<-inv.logit(intercept + linear * x  + polynomial * x^2)
   return(p)
}

predframe<-list()
for (x in 1:length(parall)){
  #correct level of effects
  f<-fixf[[x]]
  
  #split by estimate
  sf<-split(f,f$Var2)
  
  sfvalue<-lapply(sf,function(y){
    y=trajF(y$value[1],y$value[2],y$value[3],sdat[[x]]$Phylo.Relatedness)
    data.frame(x=sdat[[x]]$Phylo.Relatedness,y)
  })
  
  #bind 
  sfvalue<-melt(sfvalue,id.vars=c("x","y"))
  
  sfvalue$P_A<-sdat[[x]]$P_A
  
  colnames(sfvalue)<-c("x","y","line","P_A")
  predframe[[x]]<-sfvalue
  
  }
names(predframe)<-names(sdat)
```

```{r,fig.height=10,fig.width=14,cache=FALSE}  
#plot predicted values
mpred<-melt(predframe,id.vars=colnames(predframe[[1]]))

#order the factors
mpred$Assemblage<-factor(mpred$L1,levels=names(sdat)[c(2,1,3,4,5,6,7)])

#assign names
mpred$Type<-mpred$Assemblage %in% names(sdat)[1:3]
mpred$Type[mpred$Type == T]<-"Observed"
mpred$Type[!mpred$Type == "Observed"]<-"Simulation"

#simulated
psim<-ggplot() + geom_line(data=mpred[mpred$Type == "Simulation" & mpred$line=="Mean",],aes(x=x,y=y)) + geom_line(data=mpred[mpred$Type == "Simulation" & mpred$line=="Lower",],aes(x=x,y=y),linetype="dashed") + geom_line(data=mpred[mpred$Type == "Simulation" & mpred$line=="Upper",],aes(x=x,y=y),linetype="dashed") + facet_wrap(~Assemblage,scales="free") + theme_bw() + labs(x="Phylogenetic distance to the closest related species",y="Probability of occurrence")

pobs<-ggplot() + geom_line(data=mpred[mpred$Type == "Observed" & mpred$line=="Mean",],aes(x=x,y=y)) + geom_line(data=mpred[mpred$Type == "Observed" & mpred$line=="Lower",],aes(x=x,y=y),linetype="dashed") + geom_line(data=mpred[mpred$Type == "Observed" & mpred$line=="Upper",],aes(x=x,y=y),linetype="dashed") + facet_wrap(~Assemblage) + theme_bw() + labs(x="Phylogenetic distance to the closest related species",y="Probability of occurrence")

grid.arrange(psim,pobs)
```

```{r,echo=FALSE,results='hide',cache=FALSE}
jpeg("MixedEffects.svg",res=600,height=9,width=10,units="in")
grid.arrange(psim,pobs)
dev.off()  
```

This figure is almost identical to the bayesian estimate shown in the paper. The only exception is the size of the confidence intervals for the observed data. This is due to the relatively low sample size for some species and inflates the estimate. In a bayesian approach, the data poor species borrow strength from the hierarcichal inference, giving a more reasonable confidence interval. 

```{r}
save.image("Independence.Rdata") 
```
