---
title: 'Appendix C: SDM Evaluations'
author: "Ben Weinstein"
date: "Friday, November 14, 2014"
output: pdf_document
---

```{r,warning=FALSE,message=FALSE}
#Load required libraries
library(ggplot2)
library(knitr)
library(plyr)
library(dplyr)
library(reshape2)
library(GGally)
library(raster)

opts_chunk$set(warning=FALSE,message=FALSE)
opts_chunk$set(cache=FALSE,fig.width=12,echo=TRUE)

#Set git path
gitpath<-"C:/Users/Ben/Documents/Pred_Obs/"
droppath<-"C:/Users/Ben/Dropbox/"

#Where are the niche models saved?
savefolder<-"C:/Users/Ben/Dropbox/Thesis/Pred_Realized/NicheModels/0.1/"

setwd(savefolder)
```

#Model Evaluation

```{r}
#Get the ROC model evaluation from file
model_eval<-list.files("C:/Users/Ben/Dropbox/Thesis/Pred_Realized/NicheModels/0.1/",full.name=TRUE,recursive=T,pattern="ModelRocEval.csv")
```

```{r}
model_eval<-rbind.fill(lapply(model_eval,read.csv))
colnames(model_eval)<-c("Model","Species","Stat")
#Plot
ggplot(model_eval, aes(x=Species,y=Model,fill=Stat)) + geom_tile() + scale_fill_gradient("ROC",limits=c(0,1),low="blue",high="red",na.value="white") + theme(axis.text.x=element_text(angle=-90))

```

Number of Species included based on Ensemble Niche model support.

```{r}
model_thresh<-sapply(seq(.5,.95,.05),function(x){
  table(model_eval$Stat > x,model_eval$Model)["TRUE",]
})

colnames(model_thresh)<-seq(.5,.95,.05)
model_thresh<-melt(model_thresh)

names(model_thresh)<-c("Model","ROC_Threshold","Number_of_Species")

ggplot(model_thresh,aes(x=ROC_Threshold,y=Number_of_Species,col=Model)) + geom_line(size=1.5) + geom_point() + geom_text(aes(label=Number_of_Species),vjust=4,size=5) + theme_bw()

```

#Comparison of evaluation statistics

```{r}
model_eval<-list.files("C:/Users/Ben/Dropbox/Thesis/Pred_Realized/NicheModels/0.1/",full.name=TRUE,recursive=T,pattern="ModelEval.csv")

#some legacy error here:
model_eval<-rbind.fill(lapply(model_eval,read.csv))
colnames(model_eval)<-c("Metric","Species","GBM","GLM","MAXENT")
model_e<-melt(model_eval,id.var=c("Metric","Species"))

#Plot with facets
p<-ggplot(model_e, aes(x=Species,y=variable,fill=value)) + geom_tile() + facet_wrap(~Metric) + theme(axis.text.x=element_text(angle=-90))
p + scale_fill_gradient("Score",limits=c(0,1),low="blue",high="red",na.value="white")+ theme_bw() + theme(axis.text.x = element_blank()) 

#How correlated are model scores
cast_mods<-dcast(model_e, Species~...)

#GBM correlations
ggpairs(cast_mods[,c(2,5,8)],lower=list(continuous="smooth",method="lm"))


#GLM correcation
ggpairs(na.omit(cast_mods[,c(3,6,9)]),na.rm=TRUE,lower=list(continuous="smooth",method="lm"))


#Maxent Correlation
ggpairs(cast_mods[,c(4,7,10)],lower=list(continuous="smooth",method="lm"))


ggplot(model_e,aes(x=variable,y=value)) + facet_wrap(~Metric) + geom_boxplot()+ geom_smooth(method="lm") + theme_bw() + labs(y="Score",x="Model")
```

#Variable importance
```{r}
varI<-list.files("C:/Users/Ben/Dropbox/Thesis/Pred_Realized/NicheModels/0.1/",full.name=TRUE,recursive=T,pattern="VarImportance.csv")
varI<-rbind.fill(lapply(varI,read.csv))
varI<-varI[,-1]

#Melt variable for plotting
mvar<-melt(varI)
colnames(mvar)<-c("Bioclim","Species","Model","value")

#Plot variable importance across all models
ggplot(mvar, aes(x=Species,y=Bioclim,fill=value)) + geom_tile() + scale_fill_gradient(limits=c(0,1),low="blue",high="red",na.value="white") + theme(axis.text.x=element_text(angle=-90)) + facet_grid(Model ~ .) + ggtitle("Variable Importance")

```

#Sensitivity and specificity

* Evalaute threshold of habitat suitability which translates the niche model results into predicted assemblages.

* Calculate true positives (TP), true negatives (TN), false positives (fP), and false negatives (fn), as well as sensitivity (TP/TP+FN) and specificity (TN/TN+FP) using our observed assemblages compared to the predicted assemblages

* Also calcuate the mean habitat suitability for each threshold

```{r}
load(paste(droppath,"C:/Users/Ben/DropboxThesis/Pred_Realized/Assemblages/Threshold0.1/Results/PredictedRealized_.RData",sep=""))
```

```{r}
source(paste(gitpath,"SpeciesOverlapSourceFunctions.R",sep=""))

#load models to file, makes it more transferable than keeping them on disk
modr<-lapply(niche,raster,values=TRUE)

#Get the number of records per site, atleast 10 presences. 
records_site<-names(which(apply(siteXspp.raster,1,sum,na.rm=TRUE) > 5))
modlist<-modr[names(modr) %in% records_site]

#Lets go get the presence data on hummingbird distributions
  Niche_locals<-read.csv(paste(gitpath,"InputData\\MASTER_POINTLOCALITYarcmap_review.csv",sep=""))
  
  #Just take the columns you want. 
  PAdat<-Niche_locals[,colnames (Niche_locals) %in% c("RECORD_ID","SPECIES","COUNTRY","LOCALITY","LATDECDEG","LONGDECDEG","Decision","SpatialCheck","MapDecision")]
  
  #clean localities, we checked them against published range records and visually inspected them
  PAdat<-PAdat[!PAdat$LONGDECDEG==-6,]
  loc_clean<-PAdat[!PAdat$MapDecision=="REJECT",]

###################################################
thresholds<-seq(.05,.95,.1)

th<-lapply(thresholds,function(y){
all_models<-rbind_all(lapply(modlist,function(x){
    pred_realized(mod=x,thresh.suit=y,dispersal=FALSE,plots=FALSE,loc_clean=loc_clean,fold=tempdir())
}))})

for(x in 1:length(thresholds)){
 th[[x]]$Threshold<-thresholds[x] 
}

thdf<-rbind_all(th)

ss<-dplyr::group_by(thdf,Threshold) %>% dplyr::summarize(meanSuit= mean(Suitability[P_Apred==1]),TP=sum(P_Apred==1 & PA_Binary==1),TN=sum(P_Apred==1 & PA_Binary==1),FP=sum(P_Apred==1 & PA_Binary==0),FN=sum(P_Apred==0 & PA_Binary==1)) %>% dplyr::mutate(Sensitivity=TP/(TP+FN),Specificity=TN/(TN+FP),NPV=TN/TN+FN)

ssmelt<-melt(ss,id.var="Threshold")
ggplot(ssmelt,aes(x=as.factor(Threshold),y=value,col=variable)) + geom_line(aes(group=1),col="black") + facet_wrap(~variable,scales="free") + geom_point(size=3) + theme_bw() + labs(x="Threshold"                                                                                                                                                                          )
```

Because we are most interested in predicted suitable habitat, not species occupancy, what we want to do is maximize the number of true positives and true negatives, while minimizing the number of false negatives. The worst thing for the accuracy of the model would be alot of species occuring in predicted unsuitable habitat.

```{r}
ggplot(ss,aes(y=TP,x=TN,col=as.factor(Threshold))) + geom_point(aes(size=FN)) + geom_line(aes(group=1),col="black") + theme_bw()  + labs(col="Suitability Quantile Threshold")
```


#Conclusion

**The 0.05 quantile maximizes the number of true predicted presences and absences in the observed data, while minimizing the number of false negatives**