---
title: "Relatedness and Co-occurrence Simulation"
author: "Ben Weinstein"
date: "Friday, Feb 13, 2015"
output:
  html_document:
    toc: true
    number_sections: true
---

Aim
---

To compare the form of the function between distance to most closely related species and the probability of co-occurrence if the evolution of an occurrence trait had simultanous effects of trait conservatism and repulsion among closely related species.

#Simulate Phylogeny

Start with a simple balanced phylogeny with constructed branch lengths from 0 to 1. There needs to be a sufficient number of species to distinguish patterns from random, but not too many as to make the analysis huge and unwieldy. Let's begin with 16 species.

* Create Phylogeny
* Simulate branch lengths
* Compute cophenetic distance matrix

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
library(pez)
library(ape)
library(dplyr)
library(ggplot2)
library(reshape2)
library(vegan)
library(RColorBrewer)
library(gridExtra)
library(scales)
library(picante)

#Load Data
#load("Simulation.RData")

opts_chunk$set(warning=FALSE,message=FALSE,echo=FALSE)
opts_chunk$set(cache=TRUE, cache.path = 'SimulationsPez_cache/', fig.path='figure/',fig.width=11,fig.height=6)

#create phylogeny
trx<-compute.brlen(stree(n=16,type="balanced"))
plot(trx)

#cophenetic distance
ctrx<-cophenetic(trx)
ctrx<-ctrx/max(ctrx)
```

```{r,echo=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

#Simulation Parameters

Parameter | Description
------------- | -------------
tree|phylo object
scape.size|edge dimension of square landscape
g.center|strength of phylogenetic signal in species range centers
g.range|strength of phylogenetic signal in species range sizes
g.repulse|strength of phylogenetic repulsion
wd.all|niche width, larger values simulate broader range sizes
signal.center|simulate with phylosignal in range centers
signal.range|simulate with phylosignal in range size
same.range|make all range sizes equal
repulse|include phylogenetic repulsion in range centers
center.scale|adjust strength of phylogenetic attraction in range centers independent of g.center
range.scale|adjust strength of phylogenetic signal in range size independent of g.range
repulse.scale|adjust strength of phylogenetic repulsion independent of g.repulse
site.stoch.scale|adjust strength of random variation in species richness across sites
sd.center|sd in rnorm for the range centers, increase to get more variation in center values across species
sd.range|sd rnorm for the range sizes, increase to get more variation in range sizes across gradients
rho|Grafen branch adjustment of phylogenetic tree see corGrafen
th|probability threshold 10^-th above which species are considered present at a site


#Paramater choices

For the simulations the scape szie will be a 20*20 grid, with species able to occupy any part of the grid. This allows the species to have enough space to occupy their full range, without forcing overlap because not enough available cells. This value doesn't really matter, its the ratio of grid size to range size. I've chosen a relatively small number because i'm iterating twenty times in the model and it creates a relatively large dataframe (n=20,000). 

For our purposes, we care most about the parameters g.center and g. repulse. The help functions reads: 

```
The amount and type of structure is determened by the signal parameters g.center, g.range and g.repulse. Parameters are based on an Ornstein-Uhlenbeck model of evolution with stabilizing selection. Values of g=1 indicate no stabilizing selection and correspond to the Brownian motion model of evolution; **0<g<1 represents stabilizing selection**; and **g>1 corresponds to disruptive selection** where phylogenetic signal for the supplied tree is amplified. See corBlomberg. Communities are simulated along two gradients where the positions along those gradients, g.center and range sizes g.range, can exhibit phylogenetic signal. Phylogenetic attraction is simulated in the g.center parameter, while repulsion in g.repulse.
```
##View function
```{r}

print(scape)
```

```{r,echo=FALSE}
###Species Ranges

#Function to plot species ranges on to the simulated grid. Needs to specify the same threshold as the scape th function (see questions). Optional argument to color species 
 
#color ramp function
gs.pal <- colorRampPalette(c("red","blue"),space="rgb")

gridplot<-function(x,tree,threshold,opt.color=FALSE,plot=TRUE,guide=TRUE){
  dat<-melt(x$sppXs)
  
  #Binary presences
  dat$binary<-(dat$value>threshold)
  colnames(dat)<-c("Site1","Site2","Species","value","binary")
  
  #remove empty slots
  dat[dat$binary==FALSE,"Species"]<-NA
  
  #remove NA's
  dat<-dat[!is.na(dat$Species),]
  
  #merge color profile
  col.tip<-data.frame(Label=tree$tip.label,col=gs.pal(length(tree$tip.label)))
  
  #species number
  col.tip$Species<-which(col.tip$Label %in% tree$tip.label)

  dat<-merge(dat,col.tip,by="Species")
  

  if(opt.color){
      if(!guide){
  p<-ggplot(dat,aes(x=Site1,y=Site2,fill=col)) + geom_tile(alpha=.7)  + theme_bw() + labs(fill="Species") + scale_fill_identity(labels=col.tip$Label)
  
    }
        else {
  p<-ggplot(dat,aes(x=Site1,y=Site2,fill=col)) + geom_tile(alpha=.7)  + theme_bw() + labs(fill="Species") + scale_fill_identity(labels=col.tip$Label,guide="legend")
              }
              }
  

  if(plot) {p}
  return(p)
}
```


```{r,echo=FALSE}
#Co-occurrence wrapper function
#For each simulation, i'm interested in the probability of occurrence as a function of distance to species nearest to in the assemblage. 

#Let's define a wrapper function to calculate this data.frame
co_occur<-function(d,trx,ctrx){
 
#reshape data
colnames(d$Y)<-trx$tip.label
siteXsppm<-melt(d$Y)
colnames(siteXsppm)<-c("Locality","Species","P_A")

f<-split(siteXsppm,siteXsppm$Locality)

#remove the localities with only 1 species
f<-f[sapply(f,function(x) sum(x$P_A)) > 1]

if(length(f)==0){return(NA)}

#Get closest related species in each site
closest<-lapply(f,function(x){
  #species present
  pres<-x$Species[x$P_A ==1]
    dis<-ctrx[rownames(ctrx) %in% trx$tip.label,colnames(ctrx) %in% pres]
    apply(dis,1,function(y){
      as.matrix(min(y[!y==0]))
      })
  }) 

#melt each list, a bit ugly
tomerge<-lapply(closest,function(x){
 dm<-melt(as.matrix(x))[,-2]
 colnames(dm)<-c("Species","Phylo.Relatedness")
 return(dm)})

tomerge<-melt(tomerge,id.var=c("Species"))

tomerge<-dcast(tomerge,...~variable)
colnames(tomerge)<-c("Species","Locality","Phylo.Relatedness")

#need to be characters not factors
siteXsppm$Locality<-as.character(siteXsppm$Locality)
tomerge$Locality<-as.character(tomerge$Locality)

PA_phylo<-merge(siteXsppm,tomerge,by=c("Locality","Species"))
PA_phylo$P_A<-as.numeric(PA_phylo$P_A)

#remove Inf lengths, no co-occurrence
PA_phylo[!is.finite(PA_phylo$Phylo.Relatedness),"Phylo.Relatedness"]<-NA

return(PA_phylo)
}
  
```

```{r,echo=F}
sim<-function(trx,size,breadth,signal,repulse,repulseVal=1,replicates,plotR=T){
#Create a function to run any combiation of parameters and create the plot
sim.func<-function(){
  #probability of occupancy
  th=.6
  
  scape(trx, scape.size=size, g.center=1, g.range=1, g.repulse=repulseVal, wd.all=breadth, signal.center=TRUE, signal.range=TRUE, same.range=TRUE, repulse=repulse,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(replicates,sim.func(),simplify=FALSE)

th=.6 

### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x,trx,ctrx)})

#remove null lists
sim.occur<-sim.occur[!sapply(sim.occur,length)==1]

#Iteration count
for(x in 1:length(sim.occur)){
  sim.occur[[x]]$Iteration<-x
}
phy_comp<-rbind_all(sim.occur)

phy_comp$I<-cut(phy_comp$Iteration,c(seq(0,replicates,1)),labels=1:replicates)

nicheCplot<-ggplot(phy_comp,aes(x=Phylo.Relatedness,y=P_A)) + geom_smooth(aes(group=Iteration),method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage") + theme_bw()
if(plotR){return(nicheCplot)} else {return(phy_comp)}
}
```

---

#Visualizing the simulation.

The easiest way to quickly tell the phylogenetic structure of a run would be to color each species range on a scale from red to blue. Let's say i wanted to color by relatedness. One simple way to approximate this is to number the tips from 1 to n, and assume a divergent scale. This is not quite true, since the relationship is non linear, but its a good start.

For each visualization, we have one clade 'blue' and one clade 'red'.

#Simulation 1: Weak Phylogenetic Signal Followed by Repulsion in Occurrence Trait

We would expect red and blue clades to seperate, but then fragment to create the checkerboard pattern of distribution.

Using a logistic model, what would the functional form be between probability of co-occurrence and phylogenetic distane to closest related species, if we knew that both niche conservatism and competition were acting simultanesouly?

```{r,fig.height=6}
nicheWCplot<-sim(trx,size=10,breadth=100,signal=.1,repulse=T,replicates=50)
nicheWCplot<-nicheWCplot + ggtitle("Weak Phylogenetic Signal and Repulsion in Occurance Trait")
nicheWCplot
```

**Conclusion: Hypothesis supported. We see a non-linear relationship between distance to closest related species and probability of co-occurrence.**

#Simulation 2: Strong Phylogenetic Signal Followed by Repulsion in Occurrence Trait

We would expect red and blue clades to seperate, but then fragment to create the checkerboard pattern of distribution (see papers by Simberloff and Gotelli)

```{r,fig.height=6}
nicheCplot<-sim(trx,size=10,breadth=100,signal=1,repulse=T,replicates=50) + ggtitle("Strong Phylogenetic Signal and Repulsion in Occurance Trait")
nicheCplot
```

#Simulation 3: Weak Phylogenetic Signal Followed by No Repulsion in Occurrence Trait

```{r,fig.height=6}
nicheWCNRplot<-sim(trx,size=10,breadth=100,signal=.1,repulse=F,replicates=50)

nicheWCNRplot<-nicheWCNRplot + ggtitle("Weak Phylogenetic Signal and No Repulsion in Occurance Trait")
nicheWCNRplot
```

#Simulation 4: Strong Phylogenetic Signal Followed by No Repulsion in Occurrence Trait

```{r,fig.height=6}
 
nicheCNRplot<-sim(trx,size=10,breadth=100,signal=1,repulse=F,replicates=50)

nicheCNRplot<-nicheCNRplot + ggtitle("Weak Phylogenetic Signal and No Repulsion in Occurance Trait")
nicheCNRplot
```

#Comparison of functional form

```{r}
grid.arrange(nicheWCplot,nicheCplot,nicheWCNRplot,nicheCNRplotncol=2) 
```

```{r,echo=FALSE,eval=TRUE}
#Write simulations to file to be used in bayesian estimation
ps<-list(Weak=phy_comp_model2,Strong=phy_comp_model2)
for(x in 1:length(ps)){
  ps[[x]]$Hyp<-names(ps)[x]
}

nda<-lapply(ps,nrow) 

psdat<-rbind_all(ps)
write.csv(psdat,"C:/Users/Ben/Documents/Pred_Obs/Bayesian/simdats.csv")
```

#Varying levels of repulsion and signal

One important question would be the sensitivity of the level of repulsion and range signal in creating our predicted pattern.

```{r,fig.height=6}
sim.func<-function(repulse,signal,scale){scape(trx, scape.size=20, g.center=signal, g.range=signal, g.repulse=repulse, wd.all=500, signal.center=TRUE, signal.range=TRUE, same.range=TRUE, repulse=TRUE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)} 

#repeat 5 times for each set
sig<-seq(0.2,1,.2)
repel<-c(seq(0.2,1,.2))

ins<-expand.grid(sig,repel,scale=1) 
ins<- ins[rep(seq_len(nrow(ins)), 10), ]
colnames(ins)<-c("Signal","Repulse") 

#container
sim.list<-list()

for (x in 1:nrow(ins)){
  sim.list[[x]]<-sim.func(signal=ins[x,1],repulse=ins[x,2])
}

### Compute co-occurrence
sim.occur<-list()
for(x in 1:length(sim.list)){
  phy<-co_occur(sim.list[[x]],trx,ctrx)
  phy$Repulsion<-ins[x,"Repulse"]
  phy$Signal<-ins[x,"Signal"]
  sim.occur[[x]]<-phy
  }

sim.occur<-sim.occur[!sapply(sim.occur,length)==3]
 
phy<-rbind_all(sim.occur)
ggplot(phy,aes(x=Phylo.Relatedness,y=P_A,col=Repulsion)) + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2))),aes(group=as.factor(Repulsion)),se=F) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage") + theme_bw() + facet_wrap(~Signal,scales="free") + labs(col="Repulsion") + scale_color_continuous(low="blue",high="red")
```

Each facet is a value of g.center (phylogenetic signal) for the occurrence trait, and each color is a value of repulsion among closely related species. Fifty replicates were used for each run.
Conclusion: The observed pattern is pretty robust to the exact parameters of g.center (facets) and g.repulse (color). 

#Does # of tips effect the pattern?

Let's run a series of tree shapes and just look at the logistic function.

```{r}
#create Phylogeny
trees<-function(sp,type,size,breadth,replicates){

#create phylogeny
trx<-compute.brlen(stree(n=sp,type))

#cophenetic distance
ctrx<-cophenetic(trx)
ctrx<-ctrx/max(ctrx)

#Threshold for presence
th=.6
#Compute Models
nicheWCplot<-sim(trx,size=size,breadth=breadth,signal=.1,repulse=T,replicates=replicates)

nicheCplot<-sim(trx,size=size,breadth=breadth,signal=1,repulse=T,replicates=replicates)

nicheWCNRplot<-sim(trx,size=size,breadth=breadth,signal=.1,repulse=F,replicates=replicates)

nicheCNRplot<-sim(trx,size=size,breadth=breadth,signal=1,repulse=F,replicates=replicates)

grid.arrange(nicheWCplot,nicheCplot,nicheWCNRplot,nicheCNRplot,ncol=2)
}
```

##Balanced Trees of 8,16,32 Species

### 8 Tips
```{r,fig.height=5}
trees(8,"balanced",size=10,breadth=100,replicates=2) 
```

### 16 Tips

```{r,fig.height=5}
trees(16,"balanced",repulse=1,size=20,breadth=500) 
```

### 32 Tips
```{r,fig.height=5}
trees(32,"balanced",repulse=1,size=20,breadth=500)  
```


##Unbalanced Trees of 8,16,32 Species

### 8 Tips
```{r,fig.height=5}
trees(8,"balanced",repulse=1,size=20,breadth=500) 
```

### 16 Tips

```{r,fig.height=5}
trees(16,"balanced",repulse=1,size=20,breadth=500) 
```

### 32 Tips
```{r,fig.height=5}
trees(32,"balanced",repulse=1,size=20,breadth=500)  
```

#Conclusion
The polynomial pattern seems robust to tree shape. It seems dulled by increasing tree size, i.e that it takes more repulsion in larger trees to decrease co-occurrance among closely related species. The polynomial pattern must be the balance between the trait conservatism and the repulsion, and with more species its probably a narrower range that creates this pattern.


### As # of tips increases, there is more stochasticity

##64 Tips

```{r}
trx<-compute.brlen(stree(64,"balanced"))
ctrx<-cophenetic(trx)

#Model 3

sim.list<-replicate(10,sim.func(),simplify=FALSE)

th=.6 
### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x,trx,ctrx)})

#remove null lists
sim.occur<-sim.occur[!sapply(sim.occur,length)==1]

gridlist<-lapply(sim.list[1:6],function(x){
  gridplot(x,tree=trx,threshold=-log10(th),opt.color=TRUE,guide=FALSE)})

multiplot(plotlist=gridlist,cols=3)

#Iteration count
for(x in 1:length(sim.occur)){
  sim.occur[[x]]$Iteration<-x
}
phy_comp<-rbind_all(sim.occur)

phy_comp$I<-cut(phy_comp$Iteration,c(seq(0,10,1)),labels=1:10)

nicheWCplot<-ggplot(phy_comp,aes(x=Phylo.Relatedness,y=P_A)) + geom_smooth(aes(group=I),method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage") + ggtitle("Weak Phylogenetic Signal and Strong Repulsion in Occurance Trait") + theme_bw()

nicheWCplot
```


```{r}
save.image("Simulation.RData")
````