---
title: "Relatedness and Co-occurrence Simulation"
author: "Ben Weinstein"
date: "Friday, October 24, 2014"
output:
  html_document:
    toc: true
    number_sections: true
---

**Disclaimer: The pez::scape function is a rewritten picante::pglmm.sim with much better user input written by Matt Helmus. This work is unpublished and is not to be distributed without explicit consent of package creator Will Pearse. Many thanks to Will and Matt Helmus for their work pushing the field forward.**

Aim
---
To describe the form of the function between distance to most closly related species and probability of co-occurrence under known mechanisms of assembly. I hypothesis that a this relationship is non-linear (2nd polynomial), such that species should increase in co-occurrence due to niche conservatism, then decrease in co-occurrence based on competition. I will then compare this simulated data with known mechanism to a empirical data using Andean Hummingbirds.

#Simulate Phylogeny

Start with a simple balanced phylogeny with branch lengths

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
library(pez)
library(ape)
library(dplyr)
library(ggplot2)
library(reshape2)
library(phytools)
library(vegan)
library(RColorBrewer)
library(gridExtra)
library(scales)

opts_chunk$set(warning=FALSE,message=FALSE,echo=TRUE)
opts_chunk$set(cache=TRUE, cache.path = 'SimulationsPez_cache/', fig.path='figure/',fig.width=10)

#create Phylogeny
trx<-compute.brlen(stree(n=16,type="left"))
plot(trx)

#cophenetic distance
ctrx<-cophenetic(trx)
ctrx<-ctrx/max(ctrx)
```

```{r,echo=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

#Simulation Parameters

Since this is a relatively unexplored landscape, let's assign a variable to all parameters and we can explore them.

Parameter | Description
------------- | -------------
tree|phylo object
scape.size|edge dimension of square landscape
g.center|strength of phylogenetic signal in species range centers
g.range|strength of phylogenetic signal in species range sizes
g.repulse|strength of phylogenetic repulsion
wd.all|niche width, larger values simulate broader range sizes
signal.center|simulate with phylosignal in range centers
signal.range|simulate with phylosignal in range size
same.range|make all range sizes equal
repulse|include phylogenetic repulsion in range centers
center.scale|adjust strength of phylogenetic attraction in range centers independent of g.center
range.scale|adjust strength of phylogenetic signal in range size independent of g.range
repulse.scale|adjust strength of phylogenetic repulsion independent of g.repulse
site.stoch.scale|adjust strength of random variation in species richness across sites
sd.center|sd in rnorm for the range centers, increase to get more variation in center values across species
sd.range|sd rnorm for the range sizes, increase to get more variation in range sizes across gradients
rho|Grafen branch adjustment of phylogenetic tree see corGrafen
th|probability threshold 10^-th above which species are considered present at a site


#Paramater choices

For the simulations the scape szie will be a 10*10 grid, with species ranges equal to 20 cells. This allows the species to have enough space to occupy their full range, without forcing overlap because not enough available cells. This value doesn't really matter, its the ratio of grid size to range size. I've chosen a relatively small number because i'm iterating twenty times in the model and it creates a relatively large dataframe (n=20,000) to fit the bayesian model.

##View function
```{r}
print(scape)
```

```{r,fig.width=10}
probPlot<-function(data,col,facet){
X1<-melt(data$X1)
X1$Gradient<-"1"
X2<-melt(data$X2)
X2$Gradient<-"2"
colnames(X1)<-c("Site","Species","Prob","Gradient")
colnames(X2)<-c("Site","Species","Prob","Gradient")
f<-rbind(X1,X2)

p<-ggplot(f,aes_string(x="Site",col=col,y="Prob")) + geom_line() + theme_bw()
p
p<-p+ facet_wrap(as.formula(facet))
p + labs(col=col,y="Probability of Presence")
}

sim.dat<-scape(trx, scape.size=20, g.center=1, g.range=1, g.repulse=1, wd.all=30, signal.center=TRUE, signal.range=FALSE, same.range=FALSE, repulse=FALSE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=3, sd.range=1,rho=NULL, th=5)

probPlot(sim.dat,col="Gradient",facet="~Species")
```


```{r}

###Species Ranges

#Function to plot species ranges on to the simulated grid. Needs to specify the same threshold as the scape th function (see questions). Optional argument to color species 

#color ramp function
gs.pal <- colorRampPalette(c("red","blue"),space="rgb")

gridplot<-function(x,tree,threshold,opt.color=FALSE,plot=TRUE,guide=TRUE){
  dat<-melt(x$sppXs)
  
  #Binary presences
  dat$binary<-(dat$value>threshold)
  colnames(dat)<-c("Site1","Site2","Species","value","binary")
  
  #remove empty slots
  dat[dat$binary==FALSE,"Species"]<-NA
  
  #remove NA's
  dat<-dat[!is.na(dat$Species),]
  
  #merge color profile
  col.tip<-data.frame(Label=tree$tip.label,col=gs.pal(length(tree$tip.label)))
  
  #species number
  col.tip$Species<-which(col.tip$Label %in% tree$tip.label)

  dat<-merge(dat,col.tip,by="Species")
  

  if(opt.color){
      if(!guide){
  p<-ggplot(dat,aes(x=Site1,y=Site2,fill=col)) + geom_tile(alpha=.7)  + theme_bw() + labs(fill="Species") + scale_fill_identity(labels=col.tip$Label)
  
    }
        else {
  p<-ggplot(dat,aes(x=Site1,y=Site2,fill=col)) + geom_tile(alpha=.7)  + theme_bw() + labs(fill="Species") + scale_fill_identity(labels=col.tip$Label,guide="legend")
              }
              }
  

  if(plot) {p}
  return(p)
}
```

Let's say i wanted to color by relatedness. One simple way to approximate this is to number the tips from 1 to n, and assume a divergent scale. This is not quite true, since the relationship is non linear, but its a start. 

**From now one, we have one clade 'blue' and one clade 'red'.**


```{r}
#Co-occurrence wrapper function
#For each simulation, i'm interested in the probability of occurrence as a function of distance to species nearest to in the assemblage. 

#Let's define a wrapper function to calculate this data.frame
co_occur<-function(d){

#reshape data
colnames(d$Y)<-trx$tip.label
siteXsppm<-melt(d$Y)
colnames(siteXsppm)<-c("Locality","Species","P_A")

f<-split(siteXsppm,siteXsppm$Locality)

#remove the localities with only 1 species
f<-f[sapply(f,function(x) sum(x$P_A)) > 1]

if(length(f)==0){return(NA)}

#Get closest related species in each site
closest<-lapply(f,function(x){
  #species present
  pres<-x$Species[x$P_A ==1]
    dis<-ctrx[rownames(ctrx) %in% trx$tip.label,colnames(ctrx) %in% pres]
    apply(dis,1,function(y){
      as.matrix(min(y[!y==0]))
      })
  }) 

#melt each list, a bit ugly
tomerge<-lapply(closest,function(x){
 dm<-melt(as.matrix(x))[,-2]
 colnames(dm)<-c("Species","Phylo.Relatedness")
 return(dm)})

tomerge<-melt(tomerge,id.var=c("Species"))

tomerge<-dcast(tomerge,...~variable)
colnames(tomerge)<-c("Species","Locality","Phylo.Relatedness")

#need to be characters not factors
siteXsppm$Locality<-as.character(siteXsppm$Locality)
tomerge$Locality<-as.character(tomerge$Locality)

PA_phylo<-merge(siteXsppm,tomerge,by=c("Locality","Species"))
PA_phylo$P_A<-as.numeric(PA_phylo$P_A)

#remove Inf lengths, no co-occurrence
PA_phylo[!is.finite(PA_phylo$Phylo.Relatedness),"Phylo.Relatedness"]<-NA

return(PA_phylo)
}
  
```

---

#Simulation 1: Lottery Model
Assemblage structure is made solely based BM model of evolution with strong stabilizing in traits along two environmental variables.

We expect a distance to related species and occurrence to be flat, with no explanatory power.

Visually, there should be a mix of blue (clade 1) and red (clade 2). 

```{r}
#Threshold for presence
th=.6

sim.func<-function(){scape(trx, scape.size=7, g.center=1, g.range=1, g.repulse=1, wd.all=15, signal.center=FALSE, signal.range=FALSE, same.range=TRUE, repulse=FALSE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(50,sim.func(),simplify=FALSE)

gridlist<-lapply(sim.list[1:6],function(x){
  gridplot(x,tree=trx,threshold=-log10(th),opt.color=TRUE,guide=FALSE)})

multiplot(plotlist=gridlist,cols=3)

```

###Co-occurrence as a function of relatedness

```{r}
### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x)})

phy_lottery<-rbind_all(sim.occur)
ggplot(phy_lottery,aes(x=Phylo.Relatedness,y=P_A))  + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage")
```

**Conclusion: Hypothesis confirmed. ** Although the line is slightly positive, i believe this is due to the fact that there are more species are distance of .2 then there are species at distance of .1 or .8, thus more liklihood of co-occurrence.

##Try another assemblage size, grid = 10X10 and range=20 cells. 

```{r}
#Threshold for presence
th=.6

sim.func<-function(){scape(trx, scape.size=10, g.center=1, g.range=1, g.repulse=1, wd.all=20, signal.center=FALSE, signal.range=FALSE, same.range=TRUE, repulse=FALSE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(20,sim.func(),simplify=FALSE)

```

```{r}
### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x)})

phy_lottery2<-rbind_all(sim.occur)
ggplot(phy_lottery2,aes(x=Phylo.Relatedness,y=P_A))  + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage")
```

###Try much larger number of replicates.

For the bayesian model, convergence takes a very long time if fitting more than 50,000 data points. I ran 10 replicates of the lottery model for bayesian, we should get the identical pattern for 50 replicates (big data), it will just narrow the CI intervals.


```{r}
#Threshold for presence
th=.6

sim.func<-function(){scape(trx, scape.size=7, g.center=1, g.range=1, g.repulse=1, wd.all=15, signal.center=FALSE, signal.range=FALSE, same.range=TRUE, repulse=FALSE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(50,sim.func(),simplify=FALSE)

```

```{r}
### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x)})

phy_lottery2<-rbind_all(sim.occur)
ggplot(phy_lottery2,aes(x=Phylo.Relatedness,y=P_A))  + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage")
```

---

#Simulation 2: Niche conservatism and co-occurrence

We expect that species in clade blue should overlap and species in clade red should overlap, but there should be no/little red-blue overlap.

We expect that as there are a more closely related species, there should be greater probability of co-occurrence.

```{r,height=20}
#Threshold for presence
th=.6

sim.func<-function(){scape(trx, scape.size=7, g.center=1.1, g.range=1.1, g.repulse=1, wd.all=15, signal.center=TRUE, signal.range=TRUE, same.range=TRUE, repulse=FALSE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(50,sim.func(),simplify=FALSE)

gridlist<-lapply(sim.list[1:6],function(x){
  gridplot(x,tree=trx,threshold=-log10(th),opt.color=TRUE,guide=FALSE)})
multiplot(plotlist=gridlist,cols=3)
```

###Co-occurrence as a function of relatedness

```{r}
### Compute co-occurrence
sim.occur<-list()
for (x in 1:length(sim.list)){
  phy<-co_occur(sim.list[[x]])
  phy$Iteration<-x
  sim.occur[[x]]<-phy
  }

phy_nichec<-rbind_all(sim.occur)

ggplot(phy_nichec,aes(x=Phylo.Relatedness,y=P_A)) + geom_smooth(aes(group=1),method="glm",family="binomial",formula=(y~(poly(x,2))))+ labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage")
```

**Conclusion: As relatedness increases, species tend to co-occur more. This fits out prediction of niche conservatism. **

##Try another assemblage size, grid = 10X10 and range=5 cells. 

```{r,height=20}
#Threshold for presence
th=.6

sim.func<-function(){scape(trx, scape.size=10, g.center=2, g.range=2, g.repulse=1, wd.all=20, signal.center=TRUE, signal.range=TRUE, same.range=TRUE, repulse=FALSE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(20,sim.func(),simplify=FALSE)

```

```{r}
### Compute co-occurrence
sim.occur<-list()
for (x in 1:length(sim.list)){
  phy<-co_occur(sim.list[[x]])
  phy$Iteration<-x
  sim.occur[[x]]<-phy
  }

phy_nichec2<-rbind_all(sim.occur)

ggplot(phy_nichec2,aes(x=Phylo.Relatedness,y=P_A)) + geom_smooth(aes(group=1),method="glm",family="binomial",formula=(y~(poly(x,2))))+ labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage")
```

#Simulation 3: Niche Conservatism and competition

We would expect red and blue clades to seperate, but then fragment to create the checkerboard pattern of distribution (see papers by Simberloff and Gotelli)

```{r,fig.height=10}
sim.func<-function(){scape(trx, scape.size=7, g.center=1.1, g.range=1.1, g.repulse=2, wd.all=15, signal.center=TRUE, signal.range=TRUE, same.range=TRUE, repulse=TRUE,center.scale = 1, range.scale = 1, repulse.scale = 100, site.stoch.scale = .0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(50,sim.func(),simplify=FALSE)

gridlist<-lapply(sim.list[1:6],function(x){
  p<-gridplot(x,tree=trx,threshold=-log10(th),opt.color=TRUE,guide=F)})
multiplot(plotlist=gridlist,cols=3)

#gridlist<-lapply(sim.list[1:5],function(x){
#probPlot(x,col="Species",facet="~Gradient")})

#multiplot(plotlist=gridlist,cols=2) 
```

###Co-occurrence as a function of relatedness

```{r}
th=.6 
### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x)})

#remove null lists
sim.occur<-sim.occur[!sapply(sim.occur,length)==1]

phy_comp<-rbind_all(sim.occur)

ggplot(phy_comp,aes(x=Phylo.Relatedness,y=P_A)) + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage")
```

##Try another assemblage size, grid = 7X7 and range =5 cells. 

```{r,fig.height=10}
sim.func<-function(){scape(trx, scape.size=7, g.center=1.1, g.range=1.1, g.repulse=2, wd.all=5, signal.center=TRUE, signal.range=FALSE, same.range=TRUE, repulse=TRUE,center.scale = 1, range.scale = 1, repulse.scale = 100, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(20,sim.func(),simplify=FALSE)

```

```{r}
th=.6 
### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x)})

#remove null lists
sim.occur<-sim.occur[!sapply(sim.occur,length)==1]

phy_comp2<-rbind_all(sim.occur)

ggplot(phy_comp2,aes(x=Phylo.Relatedness,y=P_A)) + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage")
```

**Conclusion: Hypothesis supported. We see a non-linear relationship between distance to closest related species and probability of co-occurrence.**

#Varying levels of repulsion and signal

One important question would be the sensitivity of the level of repulsion and range signal in creating our predicted pattern.


```{r,fig.height=10}
sim.func<-function(repulse,signal){scape(trx, scape.size=7, g.center=signal, g.range=1, g.repulse=repulse, wd.all=5, signal.center=TRUE, signal.range=TRUE, same.range=TRUE, repulse=TRUE,center.scale = 1, range.scale = 1, repulse.scale = 100, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)} 

#repeat 5 times for each set
sig<-seq(1.5,5,.5)
repel<-seq(1.5,5,.5)
ins<-expand.grid(sig,repel)
ins<-rbind(ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins,ins)
colnames(ins)<-c("Signal","Repulse") 

#container
sim.list<-list()

for (x in 1:nrow(ins)){
  sim.list[[x]]<-sim.func(signal=ins[x,1],repulse=ins[x,2])
}

```

###Co-occurrence as a function of relatedness

```{r}
### Compute co-occurrence
sim.occur<-list()
for(x in 1:length(sim.list)){
  phy<-co_occur(sim.list[[x]])
  phy$Repulsion<-ins[x,"Repulse"]
  phy$Signal<-ins[x,"Signal"]
  sim.occur[[x]]<-phy
  }

sim.occur<-sim.occur[!sapply(sim.occur,length)==3]
 

phy<-rbind_all(sim.occur)
ggplot(phy,aes(x=Phylo.Relatedness,y=P_A,col=as.factor(Repulsion))) + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage") + theme_bw() + facet_wrap(~Signal,scales="free") + labs(col="Repulsion") + scale_color_brewer(palette="RdBu")
```

Conclusion: The observed pattern is pretty robust to the exact parameters of g.center (facets) and g.repulse (color). 

Write simulations to file to be used in bayesian estimation
```{r}
ps<-list(Lottery=phy_lottery,NicheC=phy_nichec,Comp=phy_comp)
for(x in 1:length(ps)){
  ps[[x]]$Hyp<-names(ps)[x]
}

lapply(ps,nrow) 

psdat<-rbind_all(ps)
write.csv(psdat,"C:/Users/Ben/Documents/Pred_Obs/Bayesian/simdats.csv")
```

#Does tree size effect the predicted pattern?

Run a series of tree sizes, 4-32 species. For this analysis, i'm using generalized linear models with binomial fits, since each species has the same occurrence probability, this would be extremely similiar to the hiericichal bayes model, and saves alot of time for sensitivity analysis.

```{r,eval=FALSE}
#create Phylogeny
treeS<-function(size){
  browser()
trx<-compute.brlen(stree(n=size,type="balanced"))
plot(trx)

#cophenetic distance
ctrx<-cophenetic(trx)

th=.6

sim.func<-function(){scape(trx, scape.size=5, g.center=10, g.range=1, g.repulse=10, wd.all=5, signal.center=TRUE, signal.range=FALSE, same.range=TRUE, repulse=TRUE,center.scale = 1, range.scale = 1, repulse.scale = 1, site.stoch.scale = 0, sd.center=1, sd.range=1,rho=NULL, th=th)}

sim.list<-replicate(20,sim.func(),simplify=FALSE)

### Compute co-occurrence
sim.occur<-lapply(sim.list,function(x){
  phy<-co_occur(x)})

#remove NA'
sim.occur2<-sim.occur[!sapply(sim.occur,function(x) length(x))==1]
if(length(sim.occur2)==0){return(NA)}

phy_comp<-rbind_all(sim.occur2)

p<-ggplot(phy_comp,aes(x=Phylo.Relatedness,y=P_A))  + geom_smooth(method="glm",family="binomial",formula=(y~(poly(x,2)))) + labs(y="Probability of Presence",x="Phylogenetic Distance to Nearest Species in the Assemblage") + ggtitle(paste("Species",size,sep="="))
return(p)
}

sizes<-c(4,8,16,32)
treeS(8)
treeS(16)
treeS(4)
pl<-lapply(sizes,treeS)
multiplot(plotlist=pl,cols=2)
```

