---
title: "Environmental Variables"
author: "Ben Weinstein"
date: "May 4, 2016"
output: html_document
---

#Comments from Subject Editor
"I would like some thinking on whether the worldclim data are accurate enough in the Andes, where we know climate differs drastically within very small spatial scale. Further, please provide information on why the used three variables well explain distribution patterns of hummingbirds in the Andes - and why not including more environmental variables."

# Do adding more environmental variables change predicted distributions?

Let's choose 5 species that encompass different environmental characteristics and slowly increase the number of environmental variables.

#Read in libraries and hummingbird data

```{r,warning=FALSE,message=FALSE}
#Load required libraries
library(reshape2)
require(ggplot2)
library(picante)
library(dismo)
library(ape)
library(doSNOW)
library(gdistance)
library(foreach)
library(boot)
library(maptools)
library(rasterVis)
library(knitr)
library(vegan)
library(gridExtra)
library(R2jags)
library(stringr)
library(dplyr)
library(scales)

opts_chunk$set(warning=FALSE,message=FALSE)
opts_chunk$set(cache=FALSE, fig.path='figure/',fig.width=10,echo=TRUE)

#Set git path
gitpath<-"C:/Users/Ben/Documents/Pred_Obs/"
droppath<-"C:/Users/Ben/Dropbox/"

setwd(gitpath)

#Source SDM script
source(paste(gitpath,"SDMupdated.R",sep=""))


#Load image if desired

```

## Choose species subset

These five species represent a reasonable selection of the environmental challenges.

* Pterophanes cyanopterus: High elevation species on both sides of the andes
* Patagona gigas: Dry valley species at higher elevations
* Heliodoxa leadbeateri: East-slope species
* Colibri coruscans: Widespread species with pronounced elevation movement
* Florisuga mellivora: Low elevation species

```{r}
inLocalities<-read.csv("InputData/MASTER_POINTLOCALITYarcmap_review.csv")
dat<-inLocalities[inLocalities$SPECIES %in% c("Pterophanes cyanopterus","Florisuga mellivora","Colibri coruscans","Heliodoxa leadbeateri","Patagona gigas"),]
```

#Niche Models

Run ensemble niche models using biomod2 at a desired cell size and subset

```{r}
cell_size=0.5
envfolder<-"C:\\Users\\Ben\\Dropbox\\Thesis\\Pred_Realized\\EnvLayers/Cropped/"
savefolder<-"C:/Users/Ben/Documents/Pred_Obs/NicheTest/Models"
```

Loop through all bioclim variables and produce models for our candidate five species. 

```{r}
el<-paste("bio_",1:19,sep="")
rl<-el[-c(1,12,15)]

#remove base set used in original model

for(x in 1:length(el)){
  #layers to use
  env=c(el[c(1,12,15)],rl[1:x])
  
  savefolder<-paste("C:/Users/Ben/Documents/Pred_Obs/NicheTest/Models",paste(rl[1:x],collapse = "_"),sep="/")
  dir.create(savefolder)
  SDM_SP(cell_size,inLocalities=dat,envfolder,env=env,savefolder = savefolder)
}
```

##Import predicted suitable habitat rasters

```{r niche raster}
#Bring in niche models from the script SDM.R, get the folder from cell size argument

#Biomod Consensus ensemble niche models
niche<-list.files(paste(savefolder,cell_size,sep="/"),pattern="ensemble.gri",full.name=T,recursive=T)

#Just from the current env predictions.
niche<-niche[grep("current",niche,value=FALSE)]

#Name the list of suitability models
names(niche)<-lapply(niche,function(path){
  str_match(path,"current_(\\w+.\\w+)_ensemble.gri")[[2]]
})
r<-lapply(niche,raster)
plot(stack(r))
```

# Bring in assemblage data

These assemblage data are independent from the point localities and will be used in model validation.

```{r}
#Bring in the assemblages, cleaned from JP
Sites<-read.csv(paste(gitpath,"InputData//Sites.csv",sep=""),row.names=1)

#Bring in spatial data for the sites
#Extract env information for each site
Sites.sp<-SpatialPointsDataFrame(Sites[,c("LongDecDeg","LatDecDeg")],Sites)

#site by species lists
siteXspp<-t(read.csv(paste(gitpath,"InputData//siteXspp.csv",sep=""),row.names=1))
```

###Merge assemblage lists with geographic distribution of niche models

```{r}
#Create Spatial Points object of the rownmaes
site.raster<-raster(niche[[1]])
res(site.raster)<-cell_size

#Create PA matrix for the raster
raster.localities<-rasterize(y=site.raster,SpatialPoints(Sites.sp),fun=function(x,...) length(x))

#Which cell is each site in?
cellSites<-raster::extract(raster.localities,Sites.sp,cellnumbers=T)

#Split cell
head(cellSites<-data.frame(Sites.sp,cellSites)[,])
splitCellSites<-split(cellSites,factor(cellSites$cells))

#How many duplicate communities are there, ie. number of assemblages per cell

siteXspp.raster<-sapply(splitCellSites,function(x){
  if(nrow(x)== 1) out<-siteXspp[,colnames(siteXspp) %in% x$Community]
  if(nrow(x)>1 ) {
    (out<-apply(siteXspp[,colnames(siteXspp) %in% x$Community],1,sum) > 0)*1
  }
  return(out)
})

#Get the xy lat long of the cells with sites in it
cellSitesXY<-xyFromCell(raster.localities,as.numeric(colnames(siteXspp.raster)),spatial=TRUE)
rownames(cellSitesXY@coords)<-colnames(siteXspp.raster)

##Important to get a grasp on data overlap between the different sources
congruence<-melt(list(Phylo=rownames(ctrx),Assemblage=rownames(siteXspp.raster),Suitability=names(niche),Clades=clades$double.))

write.csv(table(congruence),paste(droppath,"Thesis/Pred_Realized/Assemblages//DataOverlap.csv",sep=""))

#write siteXspp table
write.csv(siteXspp.raster,paste(droppath,"Thesis/Pred_Realized/Assemblages//SiteXsppraster.csv",sep="/"))

cong<-as.data.frame.array(table(congruence))

#how many are complete
completeI<-cong[names(which(apply(cong,1,sum)==4)),]
```

##Bring in Niche models

```{r,fig.height=6,width=7}
#load models to file, makes it more transferable than keeping them on disk
modr<-lapply(niche,raster,values=TRUE)

#Get the number of records per site, atleast 5 presences. 
records_site<-names(which(apply(siteXspp.raster,1,sum,na.rm=TRUE) > 5))
modlist<-modr[names(modr) %in% records_site]

print(paste(length(modlist),"Species in the analysis"))
```
