---
title: "Bayesian Phylogenetic Structure"
author: "Ben Weinstein"
date: "Tuesday, November 04, 2014"
output:
  html_document:
    toc: true
---

#Aim

Insights from comparing phylogenetic and trait community spacing has led to a renewed focus on whether interspecific competition shape species occurrence. Combining theories of niche conservatism and limiting similarity, species should co-occur more with closely related species up to a threshold of niche overlap, followed by a decrease in co-occurrence. I developed a hierarchical Bayesian approach to assess probability of co-occurrence as a function of distance to the closest related species in hummingbird assemblages from Northern South America. I tested a fully constrained, varying intercepting and varying slope and intercept model.

</br>

#Bayesian Analysis of co-occurrence and relatedness.

## General Model Formula
$$ Y_i \sim Bern(p_i) $$

$$ logit(p_i) = \alpha + \beta *x + \gamma *x^2 $$

$$ \text{p is the probability of presence at a site}$$
$$ \text{x= cophenetic distance to the most closely related species }$$

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(pez)
library(ape)
library(dplyr)
library(ggplot2)
library(reshape2)
library(vegan)
library(gridExtra)
library(R2jags)
library(boot)
library(stringr)
opts_chunk$set(warning=FALSE,message=FALSE)
opts_chunk$set(cache=TRUE, cache.path = 'BayesPhylo_cache/', fig.path='figure/',fig.width=14,echo=FALSE)

runs=10000
burns=7800
```


```{r load_data}
#Set dropbox path
droppath<-"C:/Users/Ben/Dropbox/"

#Set git path
gitpath<-"C:/Users/Ben/Documents/Pred_Obs/"

#Load image if desired
load(paste(droppath,"Thesis/Pred_Realized/Assemblages/Threshold0.05/Results/PredictedRealized_Revis.RData",sep=""))

#start with just the present sites
#head(PA_m2)

dat<-PA_m2[PA_m2$Hyp %in% "Env" & PA_m2$P_R %in% "PA_Binary",c("Species","Locality","P_A","Phylo.Relatedness")]

#remove infinites
dat<-dat[is.finite(dat$Phylo.Relatedness),]

#sample data
```

#Model I - Constrained - Single Slope and Intercept

Species are constrained to have the same slope and intercept.

$$ Y_i \sim Bern(p_i) $$

$$ logit(p_i) = \alpha + \beta *x + \gamma *x^2 $$

WinBugs code:

```{r singlemodel,echo=FALSE,eval=TRUE}
setwd("C:/Users/Ben/Documents/Pred_Obs/Bayesian")
#Source model
source("Single.R")

#print model
print.noquote(readLines("Single.R"))
```

```{r jagsmodelsingle,eval=TRUE,echo=FALSE,fig.align='center',results='hide'}
#Input Data
Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$Phylo.Relatedness)

InitStage <- function() {list(alpha=.1,beta=.1,gamma=.1)}

#Parameters to track
ParsStage <- c("alpha","beta","gamma")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

mSingle = jags(inits = InitStage,
         n.chains=nc,
         model.file="Single.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
#Remove Burnin
parsSingle<-melt(mSingle$BUGSoutput$sims.array)
colnames(parsSingle)<-c("Draw","Chain","parameter","estimate")
parsSingle<-parsSingle[!parsSingle$Draw %in% 1:burns,]

#remove deviance
parsSingle<-parsSingle[!parsSingle$parameter %in% "deviance",]
```

###Visualize posteriors

```{r,fig.height=3}
###Posterior Distributions
ggplot(parsSingle,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

###Visualize chains

```{r,fig.height=3}
###Chains
ggplot(parsSingle,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

###Output table

```{r}
parGSingle<-group_by(parsSingle,parameter)
out_single<-summarise(parGSingle,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))
print(out_single)

#save.image("BayesPhylo.Rdata")
```

### Plot curves

```{r,echo=TRUE,eval=TRUE,fig.width=6,fig.align='center'}
#Load if needed

dpars<-dcast(parsSingle,...~parameter)

#sample 1000 values
samp<-dpars[sample(1:nrow(dpars),500),]

traj<-list()
for(x in 1:nrow(samp)){
  s<-samp[x,]
  p<-inv.logit(s$alpha + s$beta* dat$Phylo.Relatedness +s$gamma * dat$Phylo.Relatedness^2)
  traj[[x]]<-data.frame(x=dat$Phylo.Relatedness,y=p,Iteration=x)
}
    traj<-rbind_all(traj)

#summarise mean and interval
mtraj<-group_by(traj,x) %>% summarise(mean=mean(y),upper=quantile(y,0.975),lower=quantile(y,0.025))

ggplot(mtraj,aes(x=x,y=mean)) + geom_line(col="red",linetype="dashed") + geom_line(aes(y=upper),col="blue",linetype="dashed") + geom_line(aes(y=lower),col="blue",linetype="dashed")+ theme_bw() + labs(x="Phylogenetic distance to closest related species in an assemblage",y="Probability of occurrence")
```

#Model II - Unconstrained - Varying Intercept

Species have differing intercepts but constrained to have the same slope. This equates to a single mechanism shaping relatedness, but accounting for differences in occupancy between species. 

$$ Y_i \sim Bern(p_i) $$

$$ logit(p_i) = \alpha_{Species[i]} + \beta *x + \gamma *x^2 $$
$$ \alpha_{Species[i]} \sim N(intercept,sigmaIntercept)$$

WinBugs code:

```{r printmodel,echo=FALSE,eval=TRUE}
setwd("C:/Users/Ben/Documents/Pred_Obs/Bayesian")
#Source model
source("SingleSlope.R")

#print model
print.noquote(readLines("SingleSlope.R"))
```

Run jags. 

```{r jagsmodelsingleslope,eval=TRUE,echo=FALSE,fig.align='center',results='hide'}
#Input Data
dat$Species<-factor(dat$Species)

Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$Phylo.Relatedness,
  Species=as.numeric(dat$Species),
  s=nlevels(dat$Species))


InitStage <- function() {list(alpha=rep(0.1,Dat$s),beta=.1,gamma=.1)}

#Parameters to track
ParsStage <- c("alpha","beta","gamma","intercept","sigmaIntercept")

#MCMC options
ni <- 20000  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

mSingleS <- jags(inits = InitStage,
         n.chains=nc,
         model.file="SingleSlope.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r remove burnin}
#Remove Burnin
parsSingleS<-melt(mSingleS$BUGSoutput$sims.array)
colnames(parsSingleS)<-c("Draw","Chain","parameter","estimate")
parsSingleS<-parsSingleS[!parsSingleS$Draw %in% 1:((ni/nt)-1000),]

#seperate alpha,betas,gamma
allparsSingleS<-parsSingleS

specSingleS<-parsSingleS[str_detect(string=parsSingleS$parameter,pattern=c("alpha")),]

#remove deviance and species level characters
parsSingleS<-parsSingleS[!str_detect(string=parsSingleS$parameter,pattern=c("alpha")) & !parsSingleS$parameter %in% "deviance" ,]
```

###Visualize posteriors

```{r,fig.height=3}
###Posterior Distributions
ggplot(parsSingleS,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

###Visualize chains

```{r,fig.height=3}
###Chains
ggplot(parsSingleS,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

###Output table

```{r}
parGSingleS<-group_by(parsSingleS,parameter)
out_slope<-summarise(parGSingleS,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))
print(out_single)

save.image("BayesPhylo.Rdata")
```

```{r,fig.height=15,fig.width=7}
specSingleS$parameter<-droplevels(specSingleS$parameter)

#merge with original table
specSingleS$index<-as.numeric(str_extract(specSingleS$parameter,"\\d+"))
di<-data.frame(index=1:nlevels(dat$Species),Species=levels(dat$Species))

specSingleS<-merge(specSingleS,di,by="index")

#ggplot(specSingleS,aes(x=Species,y=estimate)) + geom_violin(fill="dark grey") + coord_flip()  + theme_bw()
```

##Plot species models

```{r,fig.height=40}

dpars<-dcast(allparsSingleS,...~parameter)
samp<-dpars[sample(1:nrow(dpars),1000),]

alpha_plot<-function(alpha,sp){
    
  traj<-list()
  for(x in 1:nrow(samp)){
    s<-samp[x,]
    p<-inv.logit(alpha[x] + s$beta* dat$Phylo.Relatedness +s$gamma * dat$Phylo.Relatedness^2)
    traj[[x]]<-data.frame(x=dat$Phylo.Relatedness,y=p,Iteration=x)
    }
  traj<-rbind_all(traj)
  
  #summarise mean and interval
  mtraj<-group_by(traj,x) %>% summarise(mean=mean(y),upper=quantile(y,0.975),lower=quantile(y,0.025))
  
  ggplot(mtraj,aes(x=x,y=mean)) + geom_line(col="red",linetype="dashed") + geom_line(aes(y=upper),col="blue",linetype="dashed") + geom_line(aes(y=lower),col="blue",linetype="dashed")+ theme_bw() + labs(x="",y="") + ggtitle(sp)
  }

alpha_index<-which(str_detect(string=colnames(samp),pattern=c("alpha")))

pl<-list()
for(x in 1:length(alpha_index)){
  i<-alpha_index[x]
  #Species
  sp<-di[x,"Species"]
  pl[[x]]<-alpha_plot(samp[,i],sp)
}

do.call("grid.arrange", c(pl, ncol=5))
```

##Model III - Unconstrained - Varying Intercept and Slope

Species are unconstrained in both slopes and intercepts. Species each repond differently and having different occupancies, but are drawn from a hyperprior 

$$ Y_i \sim Bern(p_i) $$

$$ logit(p_i) = \alpha_{Species[i]} + \beta_{Species[i]} *x + \gamma_{Species[i]} *x^2 $$
$$ \alpha_{Species[i]} \sim N(intercept,sigmaIntercept)$$
$$ \beta_{Species[i]} \sim N(slope,sigmaSlope)$$
$$ \gamma_{Species[i]} \sim N(slope2,sigmaSlope2)$$

WinBugs code:

```{r slopesandinterceptmodel,echo=FALSE,eval=TRUE}
#Source model
source("BPhylo.R")

#print model
print.noquote(readLines("BPhylo.R"))
```

Run jags

```{r jagsmodelvarying,eval=TRUE,echo=FALSE,fig.align='center',results='hide'}

dat$Species<-factor(dat$Species)

#Input Data
Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$Phylo.Relatedness,
  Species=as.numeric(dat$Species),
  s=nlevels(dat$Species))

InitStage <- function() {list(alpha=rep(0.1,Dat$s),beta=rep(0.1,Dat$s),gamma=rep(0.1,Dat$s),intercept=0.001,tauIntercept=0.001,tauSlope=0.001,tauSlope2=0.001)}

#Parameters to track
ParsStage <- c("intercept","linear","polynomial","sigmaSlope","sigmaIntercept","sigmaSlope2","alpha","beta","gamma")

#MCMC options
ni <- 100000  # number of draws from the 
nt <- 2   #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits = InitStage,
         n.chains=nc,
         model.file="BPhylo.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
#Remove Burnin
parsUS<-melt(m$BUGSoutput$sims.array)
colnames(parsUS)<-c("Draw","Chain","parameter","estimate")
parsUS<-parsUS[!parsUS$Draw %in% 1:((ni/nt)-2000),]

#seperate alpha,betas,gamma
spec<-parsUS[str_detect(string=parsUS$parameter,pattern=c("beta|alpha|gamma")),]

#remove deviance and species level characters
pars<-parsUS[!str_detect(string=parsUS$parameter,pattern=c("beta|alpha|gamma|deviance")),]
```

###Visualize chains

```{r,cache=TRUE,fig.height=3}
###Posterior Distributions
ggplot(pars,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

###Visualize posteriors

```{r,cache=TRUE,fig.height=3}
###Chains
ggplot(pars,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

###Output table

```{r}
parG<-group_by(pars,parameter)
out_slopeintercept<-summarise(parG,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))
print(out_slopeintercept)

```

##View species posteriors

```{r}
#Get index
spec$ind<-str_extract(spec$parameter,pattern="\\d+")

#Get parameter
spec$par<-str_extract(spec$parameter,pattern="\\w+")

ggplot(spec[spec$par %in% c("alpha","beta","gamma"),],aes(x=estimate)) + geom_density(alpha=.4,aes(group=factor(ind)),fill="grey80",col="grey30") + facet_wrap(~par,scale="free") + scale_color_discrete(guide="none") + theme_bw() + labs(y="Posterior Density")

```


##View Species

###Compare estimates of gamma

```{r view slope species,fig.height=40}

spec$ind<-str_extract(spec$parameter,pattern="\\d+")

#split into species
specS<-split(spec,spec$ind)

#match index with species names
for (x in 1:length(specS)){
  specS[[x]]$Species<-levels(dat$Species)[x]
}

#plot gamma slopes

specSdf<-rbind_all(specS)

gdf<-group_by(specSdf,Species) %>% filter(str_detect(parameter,"beta")) 
ord<-gdf %>% summarize(meanE=mean(estimate)) %>% arrange(meanE)

gdf$Species<-factor(gdf$Species,levels=ord$Species)

ggplot(gdf,aes(x=Species,y=estimate)) + geom_violin(fill="grey30") + theme_bw() + coord_flip()

#merge with data on relatedness
datS<-split(dat,dat$Species)

specCast<-lapply(specS,function(y){
  dpars<-dcast(y[,-3],...~par,value.var="estimate")
  samp<-dpars
  
  #species trajectories
  traj<-list()
  #Get species relatedness subset
  R<-datS[[unique(samp$Species)]]$Phylo.Relatedness
  
  for(x in 1:nrow(samp)){
    s<-samp[x,]
    p<-inv.logit(s$alpha + s$beta*R  +s$gamma * R^2)
    traj[[x]]<-data.frame(x=R,y=p,Iteration=x)
    }
  traj<-rbind_all(traj)
  
  #summarise mean and interval
  mtraj<-group_by(traj,x) %>% summarise(mean=mean(y,na.rm=T),upper=quantile(y,0.975,na.rm=T),lower=quantile(y,0.025,na.rm=T))
})


alpha_plot<-function(alpha,sp){
  ggplot(alpha,aes(x=x,y=mean)) + geom_line(col="red",linetype="dashed") + geom_line(aes(y=upper),col="blue",linetype="dashed") + geom_line(aes(y=lower),col="blue",linetype="dashed")+ theme_bw() + labs(x="",y="") + ggtitle(sp)
  }

pl<-list()
for(x in 1:length(specCast)){
  #Species
  sp<-di[x,"Species"]
  pl[[x]]<-alpha_plot(specCast[[x]],sp)
}

do.call("grid.arrange", c(pl, ncol=5))
```

#Comparison of Model Estimates 

Group by the parameters they are trying to predict
```{r,fig.width=15,fig.height=4}
comp<-list(Unconstrained_slope=out_slopeintercept,Uncontrained_Intercept=out_single,Constrained=out_slope)

compm<-melt(comp)
compm<-dcast(compm,...~variable)

compm[compm$parameter %in% c("intercept","alpha"),"Parm"]<-c("Intercept")
compm[compm$parameter %in% c("slope","beta"),"Parm"]<-c("beta")
compm[compm$parameter %in% c("slope2","gamma"),"Parm"]<-c("gamma")

ggplot(compm[!is.na(compm$Parm),],aes(x=L1,y=mean,ymax=upper,ymin=lower)) + geom_pointrange() + facet_wrap(~Parm,scales="free") + labs(y="Model",x="Posterior Estimate")

save.image("BayesPhylo.Rdata")
```

#Inference

The observed assemblages showed an initial increase in the probability of co-occurrence with increasing relatedness, followed by decreased co-occurrence among closely related species. The posterior distributions were negative for the linear term and positive for the polynomial. Explorations of species specific responses showed that narrower confidence intervals for species with more data and our results of our observed data were evaluated against a fully constrained model and a full hierarchical model that allowed species level slopes to vary. All models had negative posteriors for the linear term and positive posteriors for the polynomial term. 

These data support the non-linear relationship betwen phylogenetic relatedness and occurrence in hummingbirds. The decrease among closely related species could be evidence of limiting similarity due to niche overlap, which will be tested against null models of assemblage structure. 