---
title: "Bayesian Phylogenetic Structure"
author: "Ben Weinstein"
date: "Tuesday, November 04, 2014"
output:
  html_document:
    toc: true
---

#Aim
To estimate probability of occurrence as a function of distance to closest related species. I have done this in a normal ML framework (below), but i have aggregate all species together. This is 'unfair' since each species has a different pattern of relatedness in the phylogeny. Comparing relatedness of 0.2 is different between brilliants (many closely related speices) and patagona (few closely related species). I still want to estimate a single model, so species slopes will be drawn in a hiericichal manner. I believe the model should be polynomial

</br>

## Model Formulation

Our ultimate goal is: 
$$ y \sim dbern(p) $$
<<<<<<< HEAD

$$ logit(p) = \alpha_{species} + \beta *x + \gamma *x^2 $$

$$ \text{where} \: \alpha \text{for each species} \sim dnorm(\mu,\tau) $$

$$ \text{p is the probability of presence at a site}$$
$$ \text{x= distance to a closely related species }$$
=======

$$ logit(p) = \alpha_{species} + \beta_{species} *x + \gamma_{species} *x^2 $$

$$ \text{where} \: \alpha, \: \beta, \: \gamma \: \text{for each species are} \sim dnorm(\mu,\tau) $$

$$ \text{p is the probability of presence at a site}$$
$$ \text{x= distance to a closely related species }$$

I will proceed slowly increasing in model complexity. 
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851

```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(pez)
library(ape)
library(dplyr)
library(ggplot2)
library(reshape2)
library(vegan)
library(gridExtra)
library(R2jags)
library(boot)
library(stringr)
opts_chunk$set(warning=FALSE,message=FALSE)
opts_chunk$set(cache=TRUE, cache.path = 'BayesPhylo_cache/', fig.path='figure/',fig.width=10,echo=FALSE)
<<<<<<< HEAD
runs=15000
burns=10000
=======
runs=10000
burns=1000
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
```

#Data

```{r load_data}

#Set dropbox path
droppath<-"C:/Users/Ben/Dropbox/"

#Set git path
gitpath<-"C:/Users/Ben/Documents/Pred_Obs/"

#Load image if desired
load(paste(droppath,"Thesis/Pred_Realized/Assemblages/Threshold0.2/Results/Run.RData",sep=""))

#start with just the present sites
#head(PA_m2)

dat<-PA_m2[PA_m2$Assemblage %in% "Observed" & PA_m2$Tree %in% "Phylo" & PA_m2$Hyp %in% "Env" & PA_m2$P_R %in% "PA_Binary",c("Species","Locality","P_A","value")]

#remove infinites
dat<-dat[is.finite(dat$value),]

<<<<<<< HEAD
d<-dat[sample(1000,1:nrow(dat)),]
#sample data
=======
#sample data
dat<-dat[sample(1:nrow(dat),1000),]
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
kable(head(dat),row.names=F)
```

##Model Fitting - Single Slope and Intercept

$$ y \sim dbern(p) $$

$$ logit(p) = \alpha + \beta *x + \gamma *x^2 $$

```{r singlemodel,echo=FALSE,eval=FALSE}
setwd("C:/Users/Ben/Documents/Pred_Obs/Bayesian")
#Source model
source("Single.R")

#print model
print.noquote(readLines("Single.R"))
<<<<<<< HEAD

```

```{r jagsmodelsingle,eval=TRUE,echo=FALSE,fig.align='center'}
#Input Data
Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$value,
  Species=as.numeric(dat$Species),
  s=nlevels(dat$Species))

InitStage <- function() {list(alpha=.1,beta=.1,gamma=.1)}

#Parameters to track
ParsStage <- c("alpha","beta","gamma")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits = InitStage,
         n.chains=nc,
         model.file="Single.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
#Remove Burnin
pars<-melt(m$BUGSoutput$sims.array)
colnames(pars)<-c("Draw","Chain","parameter","estimate")
pars<-pars[!pars$Draw %in% 1:burns,]

#remove deviance
pars<-pars[!pars$parameter %in% "deviance",]
```

```{r}
###Posterior Distributions
ggplot(pars,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

```{r}
###Chains
ggplot(pars,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

=======

```

```{r jagsmodelsingle,eval=TRUE,echo=FALSE,fig.align='center'}
#Input Data
Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$value,
  Species=as.numeric(dat$Species),
  s=nlevels(dat$Species))

InitStage <- function() {list(alpha=.1,beta=.1,gamma=.1)}

#Parameters to track
ParsStage <- c("alpha","beta","gamma")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits = InitStage,
         n.chains=nc,
         model.file="Single.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
#Remove Burnin
pars<-melt(m$BUGSoutput$sims.array)
colnames(pars)<-c("Draw","Chain","parameter","estimate")
pars<-pars[!pars$Draw %in% 1:burns,]

#remove deviance
pars<-pars[!pars$parameter %in% "deviance",]
```

```{r}
###Posterior Distributions
ggplot(pars,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

```{r}
###Chains
ggplot(pars,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

###Output table

```{r}
parG<-group_by(pars,parameter)
out<-summarise(parG,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))
print(out)

#save.image("BayesPhylo.Rdata")
```

##Random deviates
```{r,echo=TRUE,eval=TRUE,fig.width=6,fig.align='center'}
#Load if needed

dpars<-dcast(pars,...~parameter)

#sample 1000 values
samp<-dpars[sample(1:nrow(dpars),1000),]

traj<-list()
for(x in 1:nrow(samp)){
  s<-samp[x,]
  p<-inv.logit(s$alpha + s$beta* dat$value +s$gamma * dat$value^2)
  traj[[x]]<-data.frame(x=dat$value,y=p,Iteration=x)
}
    traj<-rbind_all(traj)

#summarise mean and interval
mtraj<-group_by(traj,x) %>% summarise(mean=mean(y),upper=quantile(y,0.975),lower=quantile(y,0.025))

ggplot(mtraj,aes(x=x,y=mean)) + geom_line(col="red",linetype="dashed") + geom_line(aes(y=upper),col="blue",linetype="dashed") + geom_line(aes(y=lower),col="blue",linetype="dashed")+ theme_bw() + labs(x="Phylogenetic distance to closest related species in an assemblage",y="Probability of occurrence")
```

##Model Fitting - Varying Intercept

$$ y \sim dbern(p) $$

$$ logit(p) = \alpha_{Species} + \beta *x + \gamma *x^2 $$
$$ \alpha_{Species} \sim N(intercept,sigmaIntercept)$$

```{r printmodel,echo=FALSE,eval=FALSE}
setwd("C:/Users/Ben/Documents/Pred_Obs/Bayesian")
#Source model
source("SingleSlope.R")

#print model
print.noquote(readLines("SingleSlope.R"))

```

```{r jagsmodelsingleslope,eval=TRUE,echo=FALSE,fig.align='center'}
#Input Data
Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$value,
  Species=as.numeric(dat$Species),
  s=nlevels(dat$Species))


InitStage <- function() {list(alpha=rep(0.1,Dat$s),beta=.1,gamma=.1)}

#Parameters to track
ParsStage <- c("alpha","beta","gamma","intercept","sigmaIntercept")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits = InitStage,
         n.chains=nc,
         model.file="BPhylo.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
#Remove Burnin
pars<-melt(m$BUGSoutput$sims.array)
colnames(pars)<-c("Draw","Chain","parameter","estimate")
pars<-pars[!pars$Draw %in% 1:burns,]

#seperate alpha,betas,gamma
spec<-pars[str_detect(string=pars$parameter,pattern=c("alpha")),]

#remove deviance and species level characters
pars<-pars[!str_detect(string=pars$parameter,pattern=c("alpha")),]
```

```{r}
###Posterior Distributions
ggplot(pars,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

```{r}
###Chains

ggplot(pars,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
###Output table

```{r}
parG<-group_by(pars,parameter)
out<-summarise(parG,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))
print(out)

#save.image("BayesPhylo.Rdata")
```

##Random deviates
```{r,echo=TRUE,eval=TRUE,fig.width=6,fig.align='center'}
#Load if needed

dpars<-dcast(pars,...~parameter)

#sample 1000 values
samp<-dpars[sample(1:nrow(dpars),1000),]

traj<-list()
for(x in 1:nrow(samp)){
  s<-samp[x,]
<<<<<<< HEAD
  p<-inv.logit(s$alpha + s$beta* dat$value +s$gamma * dat$value^2)
=======
  p<-inv.logit(rnorm(1,mean=s$intercept,sd=s$sigmaIntercept) + s$beta* dat$value +s$gamma * dat$value^2)
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
  traj[[x]]<-data.frame(x=dat$value,y=p,Iteration=x)
}
    traj<-rbind_all(traj)

#summarise mean and interval
mtraj<-group_by(traj,x) %>% summarise(mean=mean(y),upper=quantile(y,0.975),lower=quantile(y,0.025))

ggplot(mtraj,aes(x=x,y=mean)) + geom_line(col="red",linetype="dashed") + geom_line(aes(y=upper),col="blue",linetype="dashed") + geom_line(aes(y=lower),col="blue",linetype="dashed")+ theme_bw() + labs(x="Phylogenetic distance to closest related species in an assemblage",y="Probability of occurrence")
```

<<<<<<< HEAD
##Model Fitting - Varying Intercept

$$ y \sim dbern(p) $$

$$ logit(p) = \alpha_{Species} + \beta *x + \gamma *x^2 $$
$$ \alpha_{Species} \sim N(intercept,sigmaIntercept)$$

```{r printmodel,echo=FALSE,eval=FALSE}
setwd("C:/Users/Ben/Documents/Pred_Obs/Bayesian")
#Source model
source("SingleSlope.R")

#print model
print.noquote(readLines("SingleSlope.R"))

```

```{r jagsmodelsingleslope,eval=TRUE,echo=FALSE,fig.align='center'}
#Input Data
Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$value,
  Species=as.numeric(dat$Species),
  s=nlevels(dat$Species))


InitStage <- function() {list(alpha=rep(0.1,Dat$s),beta=.1,gamma=.1)}

#Parameters to track
ParsStage <- c("alpha","beta","gamma","intercept","sigmaIntercept")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits = InitStage,
         n.chains=nc,
         model.file="BPhylo.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
#Remove Burnin
pars<-melt(m$BUGSoutput$sims.array)
colnames(pars)<-c("Draw","Chain","parameter","estimate")
pars<-pars[!pars$Draw %in% 1:burns,]

#remove deviance and species level characters
parsplot<-pars[!str_detect(string=pars$parameter,pattern=c("alpha")) & !pars$parameter %in% "deviance",]
```

```{r}
###Posterior Distributions
ggplot(parsplot,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

```{r}
###Chains
ggplot(parsplot,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

###Output table

```{r}
parG<-group_by(parsplot,parameter)
out<-summarise(parG,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))
print(out)

#save.image("BayesPhylo.Rdata")
```

=======
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
##Model Fitting - Varying Intercept and Slope

$$ y \sim dbern(p) $$

$$ logit(p) = \alpha_{Species} + \beta_{Species} *x + \gamma_{Species} *x^2 $$
$$ \alpha_{Species} \sim N(intercept,sigmaIntercept)$$
$$ \beta_{Species} \sim N(slope,sigmaSlope)$$
$$ \gamma_{Species} \sim N(slope2,sigmaSlope2)$$


```{r slopesandinterceptmodel,echo=FALSE,eval=TRUE}
load("C:/Users/Ben/Documents/Pred_Obs/Bayesian/BayesPhylo.Rdata")

setwd("C:/Users/Ben/Documents/Pred_Obs/Bayesian")
#Source model
source("BPhylo.R")

#print model
print.noquote(readLines("BPhylo.R"))

```

<<<<<<< HEAD
```{r jagsmodelvarying,eval=TRUE,echo=FALSE,fig.align='center'}
=======
```{r jagsmodelvarying,eval=FALSE,echo=FALSE,fig.align='center'}
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
#Input Data
Dat <- list(
  y=as.numeric(dat$P_A),
  dist=dat$value,
  Species=as.numeric(dat$Species),
  s=nlevels(dat$Species))

InitStage <- function() {list(alpha=rep(0.1,Dat$s),beta=rep(0.1,Dat$s),gamma=rep(0.1,Dat$s),intercept=0.001,tauIntercept=0.001,slope=0.1,tauSlope=0.001,tauSlope2=0.001,slope2=0.1)}

#Parameters to track
ParsStage <- c("intercept","slope","sigmaSlope","sigmaIntercept","slope2","sigmaSlope2","alpha","beta","gamma")

#MCMC options
<<<<<<< HEAD
ni <- runs  # number of draws from the posterior
=======
ni <- 100000  # number of draws from the posterior
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits = InitStage,
         n.chains=nc,
         model.file="BPhylo.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)
```

```{r}
#Remove Burnin
pars<-melt(m$BUGSoutput$sims.array)
colnames(pars)<-c("Draw","Chain","parameter","estimate")
<<<<<<< HEAD
pars<-pars[!pars$Draw %in% 1:burns,]
=======
pars<-pars[!pars$Draw %in% 1:10000,]
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851

#seperate alpha,betas,gamma
spec<-pars[str_detect(string=pars$parameter,pattern=c("beta|alpha|gamma|deviance")),]

#remove deviance and species level characters
pars<-pars[!str_detect(string=pars$parameter,pattern=c("beta|alpha|gamma|deviance")),]

```

```{r,cache=FALSE}
###Posterior Distributions

ggplot(pars,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 
```

```{r,cache=FALSE}
###Chains

ggplot(pars,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
```

###Output table

```{r}
parG<-group_by(pars,parameter)
out<-summarise(parG,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))
print(out)

#save.image("BayesPhylo.Rdata")
```

#Fitting models to full dataset

Define a function that takes in species, presence/absence and distance to the closest related species for each type of assemblage, and returns the posterior estimates of X and X^2.

```{r singlemodel,echo=FALSE,eval=FALSE}
setwd("C:/Users/Ben/Documents/Pred_Obs/Bayesian")

Bayes<-function(presence,distance,species,runs,burn,plots,folder){

dir.create(folder)

#Source model
source("SingleSlope.R")

#make species a factor
species<-as.factor(species)

#Input Data
Dat <- list(
  y=as.numeric(presence),
  dist=distance,
  Species=as.numeric(species),
  s=nlevels(species))

InitStage <- function() {list(alpha=rep(.1,Dat$s),beta=.1,gamma=.1)}

#Parameters to track
ParsStage <- c("alpha","beta","gamma")

#MCMC options
ni <- runs  # number of draws from the posterior
nt <- 1    #thinning rate
nb <- 0  # number to discard for burn-in
nc <- 2  # number of chains

#Jags

m = jags(inits = InitStage,
         n.chains=nc,
         model.file="SingleSlope.jags",
         working.directory=getwd(),
         data=Dat,
         parameters.to.save=ParsStage,
         n.thin=nt,
         n.iter=ni,
         n.burnin=nb,
         DIC=T)

#Remove Burnin
pars<-melt(m$BUGSoutput$sims.array)
colnames(pars)<-c("Draw","Chain","parameter","estimate")
pars<-pars[!pars$Draw %in% 1:burn,]

#remove deviance
pars<-pars[!pars$parameter %in% "deviance",]
pars<-pars[!str_detect(string=pars$parameter,pattern=c("alpha")),]

###Posterior Distributions
ggplot(pars,aes(x=estimate)) + geom_histogram() + ggtitle("Estimate of parameters") + facet_wrap(~parameter,scale="free") + theme_bw() 

if(plots){ggsave(paste(folder,"Posteriors.jpeg",sep="/"),height=5,width=10,dpi=300)}

###Chains
ggplot(pars,aes(x=Draw,col=as.factor(Chain),y=estimate)) + geom_line() + facet_wrap(~parameter,scale="free") + theme_bw() + labs(col="Chain")
if(plots){ggsave(paste(folder,"Chains.jpeg",sep="/"),height=5,width=10,dpi=300)}

parG<-group_by(pars,parameter)
out<-summarise(parG,mean=mean(estimate),lower=quantile(estimate,0.025),upper=quantile(estimate,0.975),sd=sd(estimate))

write.csv(out,paste(folder,"parameterestimates.csv",sep="/"))

return(pars)}
```

#Test Function

```{r}
test<-Bayes(presence = dat$P_A,distance = dat$value,species=dat$Species,runs = 10000,burn=5000,plots = TRUE,folder = "test")
```

#Original data

#Read in simulated data

```{r}
simall<-read.csv("C:/Users/Ben/Documents/Pred_Obs/Bayesian/simdats.csv",row.names=1)

colnames(simall)<-c("Locality","Species","P_A","value","Hyp","Iteration","P_R")

```

```{r}
datF<-PA_m2[PA_m2$Assemblage %in% "Observed" & PA_m2$Tree %in% "Phylo",c("Species","Locality","P_A","value","P_R","Hyp")]

datF$Locality<-as.numeric(as.character(datF$Locality))
#bind with simulated data

datF<-rbind_all(list(datF,simall))

#split data into types of assemblages
#drop Env + Dispersal for observed, its just a data subset
sdat<-split(datF,list(datF$P_R,datF$Hyp),drop=TRUE)[-4]

#for now grab 10000 random samples and remove inf values
sdat<-lapply(sdat,function(x){
  x<-x[is.finite(x$value),]
  if(nrow(x) > 5000){
    x<-x[sample(1:nrow(x),nrow(x)*.5),]
  }
  return(x)
})

```


#Bayes function for each assemblage type
```{r}

parall<-lapply(sdat,FUN=function(x){
  #create folder name
  fold<-paste(unique(x$P_R),unique(x$Hyp),sep="_")
  
  test<-Bayes(presence = x$P_A,distance = x$value,species=x$Species,runs = 20000,burn=15000,plots = TRUE,folder = fold)
})
```

###combine predicted and observed outputs
```{r}

parmelt<-melt(parall,id.vars=colnames(parall[[1]]))

#add in observed predicted and eventually simualted

parmelt[parmelt$L1 %in% "P_Apred.Env","Assemblage"]<-"Predicted"

parmelt[parmelt$L1 %in% "P_Apred.Env_Dispersal","Assemblage"]<-"Predicted"

parmelt[parmelt$L1 %in% "PA_Binary.Env","Assemblage"]<-"Observed"

parmelt[parmelt$L1 %in% c("Simulated.Lottery","Simulated.Niche Conservatism","Simulated.Niche Conservatism and Competition"),"Assemblage"]<-"Simulation"

```

#Format final figure

```{r}

simulated<-ggplot(parmelt[parmelt$Assemblage %in% "Simulation",],aes(x=estimate,fill=parameter)) + facet_wrap(~L1,scale="free_y") + geom_histogram() + theme_bw() + ggtitle("Simulated") + scale_fill_discrete(guide="none")

predicted<-ggplot(parmelt[parmelt$Assemblage %in% "Predicted",],aes(x=estimate,fill=parameter)) + facet_wrap(~L1,scale="free_y") + geom_histogram() + theme_bw() + ggtitle("Predicted")+ scale_fill_discrete(guide="none")

observed<-ggplot(parmelt[parmelt$Assemblage %in% "Observed",],aes(x=estimate,fill=parameter)) + facet_wrap(~L1,scale="free_y") + geom_histogram() + theme_bw() + ggtitle("Observed")

jpeg("PosteriorAll.jpeg",dpi=300,height=10,width=6)
grid.arrange(simulated,predicted,observed,nrow=3)
dev.off
```

##parameter estimates across all models
```{r}
finaltab<-group_by(parmelt,L1,Assemblage,parameter) %>% summarize(mean=mean(estimate),upper=quantile(estimate,0.975),lower=quantile(estimate,0.025))
write.csv(finaltab,"ParameterEstimates.csv")
save.image("BayesPhylo.Rdata")
```

<<<<<<< HEAD

=======
For each model draw mean response curve with 95 interval, to paste in the right corner, maybe plot over the support of the data.

##Plot representative curves
```{r}
x=newd<-seq(0,1,.01)
pred<-data.frame(x=x,p=inv.logit(1+newd*3.873471+newd^2*-14.405))
ggplot(pred,aes(x=x,y=p)) + geom_line()
head(pred)
>>>>>>> 54f2e721ce6c6e3bc9d0644d84b95a021d904851
